### 2020-05-04 code/PS904-Project.Rmd
commit

---
title: The Pen Mightier Than the Sword? Retired Flag Officers and Generals and Opinion-Editorials
  Since 2000
author: "Pete Erickson"
date: "4/22/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("here")       # file paths
library("magrittr")   # pipe operators!
library("tidyverse")  # workhorse package
library("lubridate")  # lubridate helps with the package
library("plyr")

```

First, I import the data set and using the here () function to build the file path.  
I replace all NA/s with ) in the data set.

```{r summary}
# Read in the Data
dataset <- read_csv(here("data", "MilEditorials23Apr.csv"))
# Replace NA with 0 in the data
dataset[is.na(dataset)] <- 0

# Let's view the data so we know what kind of dataframes each of the variables are.  
dataset

# Note that I may want to convert AuthSe, AuthRank, and Service to some form of numeric.  
# But I will come back to that.

#See the variable names.  
summary(dataset)
names(dataset) 
```

```{r manipulation}

# Create a Variable for Publishing Year called PubYr
adjdataset <- dataset %>% 
  mutate(
  PubYr = year(PubDate)
) %>% print()

# Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
adjdataset %>% arrange(PubYr) %>% 
  filter(PolicySocial==1 | PartyGen==1 | PartyName ==1 |PolAdminName==1 )  %>% 
  select (AuthName, ArtTitle, Source, PubYr) %>% 
  view()

# Create a Variable Called case-count that counts the number of times by publishing year an article published 
# regarding social policy or general parties or parties by name or administrations occurs
adjdataset %>% group_by(PubYr) %>% 
  mutate(
    case_count=sum(PolicySocial ==1 |PartyGen==1 | PartyName ==1 | PolAdminName==1) 
  ) %>% 
  print

# Count the Number of Articles Published By Type
adjdataset %>% tally(PolicySecurity) 
adjdataset %>% tally(PersTroopSpt)
adjdataset %>% tally(GenAdvice)
adjdataset %>% tally(PolicySocial)
adjdataset %>% tally(PartyGen)
adjdataset %>% tally(PartyName)
adjdataset %>% tally(PolAdminName)

# need to change the type of class these vectors are - to binary

```

```{r simple graphs}

```