### 2020-07-01 code/Submissionv1.Rmd
made version changes

---
title: "When the Pen Becomes a Sword: Retired US Military Officers and Op-Ed Use in the Contemporary Press"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |
  _This paper seeks to explain how retired military officers have attempted to shape US public opinion through the publication of op-ed pieces in major newspapers. I construct and analyze a data set that captures more than 300 opinion-editorial pieces published by retired military officers in The Wall Street Journal, The Washington Post, The New York Times, The Los Angeles Times, and the USA Today since 1975, and ultimately find two trends that stand out. The first is that retired military officers are increasingly writing about topics that are outside of traditional security related topics. The second trend is that retired military officers are increasingly criticizing Presidential administrations, especially that of President Donald Trump, for reasons other than disputes regarding international uses of force. Given the polarizing political climate in the United States and that a Presidential election will occur this fall, these two trends are concerning from the perspective of civil-military relations and suggest that the norm of a professional, apolitical military establishment is under significant strain._

geometry: margin = 1.15in              # margin widths
fontsize: 12pt                         # text size
bibliography: researchpaper-ps904.bib # bibliography file
biblio-style: authoryear               # citation style
urlcolor: blue                         # links to external URLs
citecolor: black                       # links to bibliography
linkcolor: magenta                     # internal figures, footnotes, etc.
indent: true                          # paragraph format
                                       # true for indented paragraphs

header-includes:
    - \usepackage{setspace}
    - \usepackage{multirow}
    - \usepackage{indentfirst}

output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    number_sections: false
    toc: false
    fig_caption: true
    citation_package: biblatex
    includes: 
      in_header: 
        - latex-preamble.tex
  bookdown::htmldocument2: default
  bookdown::word_document2:
      toc: false
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
```

# Introduction

On June 1 2020, the President of the United States, Donald Trump, after giving remarks at the Rose Garden of the White House in Washington D.C., walked to St. John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square. Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and after several minutes, returned to the White House.

In the days and weeks that followed, a full blown civil-military relations ensued [@feaver_military_2020]. Several cabinet members and uniformed military leaders sought to distance themselves from the President's threat of invoking the Insurrection Act of 1807 to use active-duty forces to quell domestic riots that had initiated following the death of George Floyd in Minneapolis. The Secretary of Defense, Mark Esper, stated that he "did not support" use of the 1807 Insurrection Act to deploy active duty military to establish order, and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told a graduating class from the National Defense University that he "should not have been [joining the President on the walk to St. John's Church]," and that his presence "created a perception of the military involved in domestic politics" [@esper_secretary_2020; @milley_official_2020]. Multiple retired American generals also spoke out against the Trump administration's response  But for those paying close attention to what these retired military officers said, it is clear that there was some noticeable variation in their remarks. While all of them expressed deep concern over the way in which the President was threatening to use the military [@brooks_dismay_2020; @dempsey_former_2020 ], some attacked the President in much more forceful and personal terms [@goldberg_james_2020; @mullen_i_2020].

This paper examines the way in which retired military officers have attempted to influence public opinion in contemporary America through the use of opinion-editorial pieces. It is an interesting question worthy of examination for at least three reasons. First, even after leaving active duty service, military officers remain part of the military institution - one that maintains considerable trust and credibility among the American public. A 2018 Pew Research Center poll determined that 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018]. The second reason this question is important is because as an institution, the military prides itself on being nonpartisan and apolitical, an ideal that is undoubtedly under significant stress in an era of increased political polarization [@layman_party_2006; @hare_polarization_2014]. Third, this question is worth exploring because it may reveal the degree to which the first two factors - military credibility and political polarization - have influenced the concept and norm of professionalism among the military as an institution. All of these factors are worthy of exploration, especially ahead of what promises to be a very spirited presidential election this coming fall.

This paper proceeds in three parts. First, I briefly review the basis upon which the norm of apolitical and nonpartisan behavior of the military as an institution is based. Here I briefly establish a basis for political polarization in contemporary America as well as the high degree of credibility the military as an institution presently enjoys. Second, I describe the analytical approach and the data used in this project. Assembling an original data set that includes more than 300 opinion pieces published by retired military officers in major American newspapers since the late 1970s, I conduct a mixed-methods approach which ultimately reveals that although the vast majority of opinion pieces published in major newspapers seem to uphold the ideal of the military as apolitical and non-partisan, this trend has been weakening in recent years. The data shows that retired military officers are increasingly writing about topics that are not immediately related to national security, as well as increasingly criticizing if not altogether attacking political administrations more often than they did in previous eras. Third, I briefly explore the implications of these results on the military as an institution, and specifically, its ability to remain professional in a highly polarized political climate. 
\singlespace
# The Non-Partisan Military Institution, Credibility, and Polarization  
\doublespace
The ideal of the US military acting as an objective, apolitical and non-partisan institution is deep. One of the strongest origins of the norm stems from the work of the political scientist, Samuel Huntington. In his 1957 work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, Huntington asserted that two distinct forces would always be in tension with one another in a liberal democracy such as the United States. The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced [@huntington_soldier_1957]. The second force was a "societal imperative," which Huntington said generated requirements for the military in that the military could never lose so many aspects of it such that it no longer represented the liberal democracy that it defended [@huntington_soldier_1957]. Huntington's point was that America needed a strong military that could win on the battlefield, but cautioned that such a military could not be so foreign in character so as to alienate the tenets of an American liberal society, and that these forces were at odds with each other. The key to resolving the tension generated by these two opposing forces was to adopt a scheme of "objective civilian control" in which political leaders established the ends and goals of policy, while military leaders developed and implemented the military plans to accomplish these goals [@huntington_soldier_1957]. To Huntington, largely separate operating spheres between the political leaders of a democracy and the military as an institution would ensure that the military remained "professional," and thus, untainted by partisan interests and concerns. In this way, Huntington posited that the American military would best meet the task of defeating security threats while also remaining apolitical.

While Huntington's theory has enjoyed tremendous staying power (Huntington is still studied and taught in many American military institutions), he is not without critics [see, for example, @feaver_civil-military_1996; @cohen_supreme_2003; @feaver_armed_2005; and @brooks_paradoxes_2020]. It would be too burdensome to unpack all of those here, but one of the earliest and richest critiques of Huntington, and perhaps one of the most relevant given today's polarizing environment, came from S.E. Finer. Authored in 1962, Finer's book, _The Man on Horseback: The Role of the Military in Politics_, argued that Huntington's concept of professionalism simply was not strong enough to keep the military out of politics [@finer_man_1962]. Finer's concern was that of military officers adopting an ethic of professionalism such that they would actually be more likely - not less likely - to become involved in political issues and concerns [@finer_man_1962, 22-23]. Finer was especially concerned that the adoption of professionalism alone would lead to military professionals believing that they knew what was best for their country, over and above that of the specific, elected civilian leaders in office. Finer warned that "the moment the military draw this distinction between nation and the government in power, they begin to invent their own private notion of the national interest, and from this it is only a skip to the constrained substitution of this view for that of the civilian government…this purported care for the national interest as defined by the military is indeed one of their main reasons for intervening [@finer_man_1962]. Finer's suggestion was that because of the perils of professionalism as espoused by Huntington, what was really needed to keep the military out of politics was a deep internalization of "the principle of civil supremacy" [@finer_man_1962]^[
Importantly, neither Huntington's concept of professionalism nor Finer's critique of the concept are arguing for blind obedience by the military to civilian orders and directions, a point that is, in the author's opinion, often missed in debates about civil-military relations. On the contrary, both Huntington and Finer make room and acknowledge that there are instances when military disobedience is not only required, but justified. However, both agree that in general terms, these situations should be exceedingly rare.]

It is not only in the academic literature that the tradition and worth of an apolitical military has been cultivated. The operational military has also been central in upholding the military's non-partisan traditions. Retired Army General Douglas MacArthur, a hero of World War Two in the Pacific who is also well known for the controversy surrounding his conduct in the Korean War and his eventual dismissal by President Truman, told the cadets at West Point in May of 1962.

\singlespacing

> **Yours is the profession of arms, the will to win, the sure knowledge that in war there is no substitute for victory, that if you lose, the Nation will be destroyed, that the very obsession of your public service must be Duty, Honor, Country. Others will debate the controversial issues, national and international, which divide men’s minds. But serene, calm, aloof, you stand as the Nation’s war guardians, as its lifeguards from the raging tides of international conflict, as its gladiators in the arena of battle. For a century and a half you have defended, guarded and protected its hallowed traditions of liberty and freedom, of right and justice. Let civilian voices argue the merits or demerits of our processes of government. Whether our strength is being sapped by deficit financing indulged in too long, by federal paternalism grown too mighty, by power groups grown too arrogant, by politics grown too corrupt, by crime grown too rampant, by morals grown too low, by taxes grown too high, by extremists grown too violent; whether our personal liberties are as firm and complete as they should be. These great national problems are not for your professional participation or military solution. Your guidepost stands out like a tenfold beacon in the night: Duty, Honor, Country** [@macarthur_duty_1962].

\doublespacing

Current military doctrine and practices likewise reinforce the notion of the military remaining apolitical. Prior to the 2016 Presidential Election, Marine Corps General Joseph Dunford, who then served as the Chairman of the Joint Chiefs of Staff, said that the basis upon which the military is able to maintain the public's trust is it's ability to remain "apolitical and neutral" [@dunford_upholding_2016]. General Martin Dempsey, in an opinion piece published in the Washington Post on July 31, 2016, specifically criticized retired Army Major General Mike Flynn and Marine Corps General John Allen for participating in the Republican and Democratic conventions, respectively, stating that, "As generals, they have an obligation to uphold our apolitical traditions. They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated. It was a mistake for them to participate as they did. It was a mistake for our presidential candidates to ask them to do so" [@dempsey_military_2016].

Although the notion of an apolitical military is the theoretical goal of the institution, the practical record remains far from perfect. From MacArthur's clash with President Truman in the early 1950's, to the so-called "Revolt of the Generals" in 2005-2006, military leaders, even those who are retired, have ventured into political territory from time to time. What we have established, though, is clear: a non-partisan military that performs its military mission well has long been the goal of both civil-military relations theorists and military officers. More importantly, this is what fosters and ultimately earns the trust and confidence of the American people.

## Political Polarization, Public Opinion, and Military Expertise

Scholars of American Politics have studied political polarization in depth, and while they have offered differing explanations for the causes of polarization, ranging from race and religion [@abramowitz_great_2018], to debates about the size and role of government [@sides_two_2015], to fierce policy disagreements over policy [@mann_its_2012], virtually all agree that polarization, if not an entirely new phenomenon, has become more pronounced in recent decades [@hare_polarization_2014; @bonica_mapping_2014]. Hare and Poole are among scholars who argue that modern American political polarization began in the 1970s as a result of redistribution along racial lines following key pieces of legislation in the 1960s [@hare_polarization_2014]. They further argue that polarization disagreement has only intensified, something that a significant portion of modern Americans agree on [@schaeffer_far_nodate].

Within this polarized political environment, what role does the military play - particularly retired military officers who possess considerable military expertise? In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992]. Zaller argued that throughout modern America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992]. Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send, in order to be understood by regular citizens who are part of the public, must be "short, simple, and highly thematic" [@zaller_nature_1992]. In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information. He is rather acknowledging that most citizens care most about the world that is not "beyond their personal experience" [@zaller_nature_1992]. His point is that elites play a significant role in helping to form and sway public opinion.

Michael Robinson further examined how elite cues are received and processed. Robinson argues that partisan groups, as a result of ongoing political polarization, are cognitively biased in distinct ways in how they receive and process information [@robinson_who_2019]. As a result, Robinson found that partisans will express confidence in the military in vastly differing ways even after neutral, non-partisan information about various types of military failure is presented to these groups [@robinson_who_2019]. If Republicans are less likely to express a reduction in confidence in the military as an institution than Democrats are, as Robinson alleges, then even accurate, unbiased information is potentially irrelevant, as these differing partisan groups are pre-programmed to selectively see a particular situation as they wish to see it [@robinson_who_2019]. His main point serves to warn us regarding confidence in the military as an institution: there may be real reasons to lose confidence in the military as a result of a particular incident, but partisan filters will prevent a reduction of confidence from actually being expressed.

Zaller's framework makes it clear that it is a perfectly reasonable assumption that retired military officers who choose to express their opinions through the form of opinion-editorial pieces constitutes a form of elite discourse. After all, retired military officers have at least twenty years of experience, and in the case of most multi-star generals or admirals, usually more than 30 and in some cases close to 40 years of military experience. Indeed, most retired senior generals and admirals have spent their professional careers solely in the military, and have extensive experience while serving in leadership positions of large organizations and, in many cases, during times of war. Retired military officers are thus elites in the field of national security, and in line with Zaller's framework, they possess the potential to significantly contribute to policy debates and thus, to influence public opinion. Yet we also know that the American public will not necessarily evaluate the opinions of these military experts objectively, but rather through pre-conceived and potentially partisan lenses [@robinson_who_2019].

# Methodological Approach, Data, and Hypotheses

Given these three parameters -- that retired military officers are part of an institution that highly values professionalism and non-partisanship, that political polarization in America has increased since at least the 1970s, and that retired military officers are elites who possess the potential to sway public opinion -- I now aim to see whether the discourse of retired military officers has changed over time, specifically since the 1970s, as polarization has worsened. To answer this question, I assemble and examine a data set that consists of opinion pieces authored by retired military officers in major US newspapers. My methodological approach involves two key steps. First, I determine if there have been shifts in the topics that retired military officers have been writing about. Second, I apply a normative lens to determine if there have been shifts in the tone and tenor of the pieces authored by these officers, particularly regarding the professional aspects of the military previously discussed. In simple terms, I want to answer the questions, what are these retired military officers writing about, and through the lens of professionalism, should they be saying what they are saying?

Using the Factiva and Nexus-Uni databases, I conduct a search that focuses keywords in the byline (author field) such as "retired military officer," "retired from the Army, Navy, Marine Corps, Air Force," "former Chairman of the Joint Chiefs of Staff," "General," "Admiral," and so forth. For sources, I confine my search to the largest US newspapers that regularly run opinion-editorial pieces, which includes  _The Wall Street Journal_ (WSJ), _The New York Times_ (NYT), _The Washington Post_ (WaPo), _The Los Angeles Times_, and _The USA Today_. I begin the search for the data of January 1, 1979 - 30 June 2020^[I also open my search to include both print and online offerings to ensure that I am capturing any opinion pieces published digitally. Most of the newspapers initiated digital offerings in the late 1990s.]. Using newspapers, and particularly these specific ones, provides some advantages as well as disadvantages. The primary advantage is that these sources have been around since the period I wish to examine, unlike many of the online only sites that have only been present in the digital age. This allows me to form a data set which covers more time. The primary disadvantage is that I am not capturing all of the opinion-pieces authored or given by retired military officers. For example, retired military officers are common on cable TV, often weighing in on topics, and I am not capturing those here. It is also true that over time,  social media has taken over as the primary means in which a majority of Americans receive their news [@martin_how_2018]. Print newspapers still exist, but overall circulation is still down today compared to earlier years, even after factoring in the expanse of digital options available^[As a means of rough comparison, estimated total daily newspaper circulation in the US in 2000 was 62 million and 54 million for weekday and weekend editions, respectively, while in 2018, the the same figures were 29 million and 31 million [@barthel_leading_2019]]. Nonetheless, I contend that the data I am able to assemble will still be informative. Publishing opinion-editorial pieces, especially in major, competitive newspapers, is a highly competitive process. For the newspapers that I examine in this study, it is not uncommon to see opinion pieces published by figures that include heads of state, members of congress, governmental figures, and leaders of industry and academia. Additionally, opinion pieces are deliberately authored for the purpose of informing and persuading the public on a salient or timely topic. They are different from interviews or commentary that may be asked by a questioner because the author must take time to put forth his or her position in writing. The data therefore allows the researcher to truly ascertain what the specific views and arguments of an author were on a given topic or policy.

Of note, when assembling the data, some authors are explicitly identified in some of their publications as retired military officers, but not in others.^[This was true of several individuals, including Alexander Haig, a retired Army General who later served as Secretary of State under Ronald Reagan;  Robert McFarlane, a retired Marine Corps Lieutenant Colonel who later served as the National Security Adviser to President Reagan, Andrew Bacevich, a retired Army Colonel who then became a prominent historian and author; and Michael Hayden, a retired Air Force General who served as director of the National Security Agency and the Central Intelligence Agency]. I conduct a separate search for all opinion pieces penned by these authors, regardless of whether they were identified as a retired military officer or not, concluding that it is best do so for coding purposes. Ultimately, the data set contains 334 opinion-editorial publications, penned by 178 different authors or author teams.^[There are several co-authored pieces. For example, where General (Retired) David Petraeus wrote one publication and another one was written by General (Retired) David Petraeus and Michael O'Hanlon, I counted these as two different authors for coding purposes.] 10 of these pieces are written unanimously, with nods by the authors to their previous service.

From a macro perspective, consider for a moment all that the date range includes with respect to the US military. Indeed, numerous and substantive events, debates, and changes have occurred in this time period for retired military officers to weigh in on. In terms of external threats and wars, this period includes the final years of the Cold War against the Soviet Union, the first war in Iraq, operations in the Balkans, and the post-9/11 wars in Iraq and Afghanistan, as well as a host of smaller but important military operations in Grenada, Panama, Somalia, and Haiti. The military has also undergone significant social and institutional changes in the time period examined. In the late 1970s, the military had ended the draft only years before following the conclusion of the Vietnam War. The debates over "Don't Ask, Don't Tell" and policies regarding women in the combat arms are also included in the data, as are numerous discussions of military budgets, weapons systems, and alliances. 
Attempting to reconcile the fact that the military as an institution prides itself on being a non-partisan and apolitical institution, that newspaper circulations (even after accounting for digital offerings) are substantially down over the past 40 years, and that the US political environment is increasingly polarized along partisan lines, I adopt the following hypotheses:

**$H_1$: The majority of op-eds published by retired military officers over the time period examine will be apolitical and non-partisan in tone.**

**$H_2$: Because overall newspaper consumption has dropped in the past 40 years, there will be fewer opinion pieces over time authored by military officers.**

**$H_3$: Given that polarization has become increasingly pronounced in the United States in recent years, there will be an increase in the number of op-eds which reveal political and partisan sentiments (indication of a drop in professionalism).**

# First Layer: Topical Analysis

In the first level of analysis, after capturing key biographical and publication information, such as author name, article title, source, publication date, and branch of service and rank of the author, I code each observation into one of several binary categories that corresponds to the overall subject matter of the publication. The goal is to create a coding scheme by subject that is sufficiently broad but also specific enough that is helpful in differentiating the overall main point of each publication. The following coding scheme is therefore devised and implemented (all variables are binary, taking on the value of 1 if the publication corresponds to the general definition of the variable, and 0 otherwise): \singlespace

* **War_Ops**- if the piece advocates for or denounces a particular course of action directly related to operational or war fighting decisions of a clear military nature. Examples include: the wisdom of troop increases, the Surge in Iraq, whether to bomb Serbia, or even the role of the prison at Guantanamo.

* **Trp_Clt**- if the piece generally showcases or highlights the work and sacrifice of members of the armed forces; to defend the work of specific members of the Armed Forces; or to call for policing up the institutional cultures of the services. Examples include: remembering our fallen soldiers on Memorial Day; improving the institutional culture of the Army or Navy; advocating on behalf of or remembering a specific flag officer.

* **Str_Adv**- if the piece is generally written to offer some sort of strategic advice for the nation, to include the wisdom of pursuing particular alliances or agreements, or adopting a specific weapons platform or military budget. Examples include: why the US needs to strengthen NATO, build ties with China, increase defense spending, or not retire the A-10 combat plane.

* **MilPSoc**- if the piece advocates for or denounces a particular social policy that _directly influences_ the armed forces. Examples include: debates over keeping or removing the "Don't Ask, Don't Tell" policy, opening up combat roles to women, the role of Transgender members of the armed forces.

* **Dom_P**- if the piece advocates for or denounces a position on a policy that _does not directly_ influence the armed forces in its day to day activities. Examples include: support for the second amendment, immigration policy, and whether confederate statues should be taken down or left up.

* **Fig_Name**- if the piece advocates for or denounces a particular position held by a political administration, or _if the main point of the piece_ is to explicitly defend or attack a political figure by name. Examples: endorsing a Presidential candidate for office, or for attacking a President or member of Congress for a decision that was made.

\doublespace
Separating opinion-editorial publications into one of several distinct categories is an imperfect process. To aid the coding process, I first identify the thesis statement of each opinion piece, which in almost every case, makes possible a reasonably clear classification of the publication. Two points are worth mentioning. First, I attempt to keep in mind the fact that opinion pieces are intended to argue for something, and thus are, to an extent, designed to be provocative. With this in mind, I did not automatically code pieces that generally advocated for or cautioned the President in the **Admin_Fig_Name** category. Second, there were some instances in which I had to choose between placing a publication into one of two categories, say, between **Warfighting_Ops** and **Strat_Advice**, but very few if any instances in which I had to choose between categories that differed greatly, such as **TroopSptCulture** and **Admin_Fig_Name**. Table 1 reveals the breakdown of the data by article type.

```{r summary, include = FALSE}
# Read in the Data
dataset <- read_csv(here("data", "MilEditorials24June.csv"))
dataset %>% print()
# Replace NA with 0 in the data for variables War_Ops thru Fig_Nme
adjdataset <- dataset %>%
  mutate_at(
    .vars = vars(War_Ops:Fig_Nme),
    .funs = ~ case_when(is.na(.) ~0, TRUE ~ 1)
) %>% mutate(
    PubYr = lubridate::year(PubDate)
) %>% view()

summary(adjdataset)
names(adjdataset)

# count the number of observations by type of binary variable and store as new dataframe
countbytype  <- adjdataset %>% count(War_Ops, Trp_Clt, Str_Adv, MilPSoc, Dom_P, Fig_Nme) 
countbytype

#count the number of times each source occurs in the data
WSJcount <- adjdataset %>% count(Source=="WSJ")
WSJcount

NYTcount <- adjdataset %>% count(Source=="NYT")
NYTcount

WaPOcount <- adjdataset %>% count(Source=="WaPO")
WaPOcount

USATodaycount <- adjdataset %>% count(Source=="USA Today")
USATodaycount

LATimescount <- adjdataset %>% count(Source=="LA Times")
LATimescount

```

```{r table, fig.align="center", include=TRUE, echo=FALSE}
countbytype %>%
  kable(format="latex", caption="Publications by Subject") %>%
  kable_styling(latex_options = "striped", font_size=10, full_width = T) %>% 
  row_spec(0, bold=T)
```

Table 1 reveals that an overwhelming majority (289 of 334, or 86.5%) of the total number of opinion pieces written since 1979 in the major newspapers examined can be classified according to one of the following three categories: **War_Ops**, **Trp_Clt**, and **Str_Adv**. It is reasonable to conclude that these op-eds written are appropriate from the perspective of civil-military relations in that they address topics which American citizens can reasonably expect retired military officers to bring their expertise to bear upon. The remaining 45 opinion pieces (13.5% of the total) fall into one of the other three categories: **MilPSoc**, **Dom_P**, and **Fig_Nme** and require further examination.  They are not necessarily inappropriate, but they likely address topics that directly address or are closely related to issues that could be considered as partisan or at least beyond appropriate military expertise. This second layer of analysis will help expose whether there is any indication that professionalism as a norm among retired military officers appears to be weakening.

\singlespace
# The Second Layer - Are Retired Military Officers Upholding Professionalism as a Norm?
\doublespace
Of the remaining observations, two are removed upon immediate inspection. One was written by an unknown author and is entitled, "Fists Raised at West Point," and was published in _The Washington Post_ on May 13, 2016. It is removed because the article is actually a compendium of two separate letters sent to the editor by two different authors, one arguing for and the other against female West Point graduates who raised their fists upon graduating, which ignited some controversy about soon-to-be Army lieutenants purportedly making a symbol associated with the "Black Lives Matter" movement [@noauthor_fists_2016]. Because the two separate letters taken together do not take a firm position on the subject, but rather show a difference in opinion about a specific event happening in 2016, I remove the observation. The second piece I remove was written by former Democratic Congressman John Murtha, who was also a retired Marine Corps officer. As a politician, the American public can and should expect him to write in the capacity of a politician, and therefore, to make partisan arguments. Not wanting to skew the remaining data, I remove his publication for the remaining analysis. Thus, the second level of analysis proceeds on a total of 43 observations, which are listed and printed in chronological publishing order, in Table 2.

\singlespace

```{r remaining data, include = FALSE}

# Show all articles and authors who published articles on Social Policy, Political Parties, and Administrations/Political Figures
rem_obs <-adjdataset %>% arrange(PubYr) %>% 
  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
  select (Author, Title, Source, PubYr, MilPSoc, Dom_P, Fig_Nme) %>% 
  view()
```

```{r table2, fig.align = "center", include=TRUE, echo=FALSE}
  rem_obs %>%
  kable(format="latex", caption="Remaining Observations", longtable=T) %>%
  kable_styling(font_size=10, latex_options="striped", repeat_header_continued=T) %>%
  column_spec(1, width="12em") %>% 
  column_spec(2, width="16em") %>% 
  column_spec(3:4, width="2.5em") %>% 
  column_spec(5:7, width="3em") %>% 
  row_spec(0, bold=T)
```

```{r code, include=FALSE, echo=FALSE}
# Create a Variable Called case-count that counts the number of times by publishing year an article published 
# regarding social policy or general parties or parties by name or administrations occurs
rem_obs %>% group_by(PubYr) %>% 
  mutate(
    case_count=sum(MilPSoc ==1 | Dom_P==1 | Fig_Nme==1) 
  ) %>% view()
```
\doublespace

A couple of interesting observations are worth noting about the data in Table 2. First, one can quickly observe that there appear to be three distinct clusters of observations with respect to publishing year. The first occurs in the time frame between roughly 1998-1999, the second between 2004-2007, and the third between 2016-2019. The clusters can also be seen in time in Figure 1 on the following page.

In the first cluster, the publications refer to multiple disagreements certain retired military officers had about how President Clinton was planning or using military force. One observation, that written by retired Marine Corps General Carl Mundy, discussed implications of changing the policy of "Don't Ask, Don't Tell," which was in place at the time, while another observation, that written by retired Army General Barry McCaffrey, refers to domestic drug policy in the United States - a perfectly acceptable topic given that he, at the time, was serving as President Clinton's Director of the Office of National Drug Control Policy. The second cluster refers to pieces penned around the so-called "Revolt of the Generals," referring to sharp disagreements that entered the public arena between multiple US generals and admirals and President George W. Bush's Secretary of Defense, Donald Rumsfeld. Disagreement about Rumsfeld's leadership, particularly regarding the planning and execution of the early years of the war in Iraq, were central during this period of time. Retired Army General Tommy Franks penned two observations that essentially provided public support for the President at the time, George W. Bush, and Retired Army Major General Paul Eaton, who was a particularly vocal critic of Rumsfeld, authored two of the pieces published in this time period. Finally, the 2016-2019 cluster is indicative mainly of concerns over the candidate and later President Donald Trump. Retired Air Force General Michael Hayden, Navy Admirals Mike Mullen and William McRaven, and Army General Stanley McChrystal are among the retired officers penning pieces during this time period.

```{r fig1, fig.height=5, fig.width=7, fig.align="center", include=TRUE, echo=FALSE}
datatograph <- data.frame(
  year=c("1994","1996","1998","1999","2000","2003", "2004", "2006", "2007", "2010", "2013","2014", "2016", "2017", "2018", "2019", "2020"), count=c(1, 1,2, 3, 1, 1, 3, 3, 3, 1, 1, 1, 6, 5, 4, 6, 1)
)
ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
  labs(
    x = "Year",
    y = "Count",
    color = NULL,
    title = "Figure 1: # of Op-Eds Written By Military Retirees",
    subtitle = "On Topics of Social Policy, Domestic Policy, or Individual Figures"
  )
```

Evaluating a conclusive level of appropriateness of the opinion pieces listed in Table 2 cannot be definitively solved, precisely because it matters the perspective from which one views how military officers should interact with the American public, especially when attempting to shape public opinion. However, it is possible to determine whether there has been deviation from non-partisan and apolitical norms. Recall that Retired Army General Douglas MacArthur cautioned the cadets at West Point in 1962 about involving themselves in matters that do not require a military solution (in MacArthur's words, of "politics grown to corrupt," "morals grown too low," and "whether our personal liberties are as firm and complete as they should be") [@macarthur_duty_1962], and that more recently, in 2016, Retired Army General Martin Dempsey lamented the participation of retired generals in the political process by endorsing political candidates and appearing at the conventions of political parties:

\singlespace

> **The American people should not wonder where their military leaders draw the line between military advice and political preference. And our nation's soldiers, sailors, airmen and Marines should not wonder about the political leanings and motivations of their leaders. As generals, they have an obligation to uphold our apolitical traditions. They have just made the task of their successors – who continue to serve in uniform and are accountable for our security - more complicated** [@dempsey_military_2016].

\doublespace

Using these ideals as guidelines allows us to conduct a second level of analysis, with emphasis on the two characteristics of topic (is the author writing about a reasonable military topic) and political leaning (is the author expressly attacking or supporting specific political figures in a political way, from which one can reasonable infer the officer's political or partisan bent)?

Using these characteristics as guidelines, I therefore develop two new variables. The first is **Result**, which can take on the following values:\singlespace
* **Permissible** -if the reader cannot clearly draw conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication in fact does in fact fall underneath the area of clear military expertise.
* **Questionable** -if the reader can draw some general, but not final, conclusions regarding the author's political stances or partisan preferences; or if the topic addressed in the publication is questionably an area of reasonable military expertise.
* **Inappropriate** -if the reader can draw clear conclusions regarding the author's partisan preferences or political leanings; or if the author addresses a topic that is not an area considered to be directly relevant to military expertise, or a matter that the military should judge.

\doublespace
The second new variable I develop is **Justification** in which I briefly comment on why I assigned the value of **Result** for the observation. Here I simply comment on whether it is a concern of partisan preferences being revealed and/or topic addressed that are responsible for the value of **Result**. The results of this second level of analysis are presented in Table 3. Note that several publications are deemed to take on values of permissible under the variable **Result**. Importantly, three of these pieces occurred during the first previously identified cluster, 1998-1999. No pieces from the second cluster, 2004-2007 have been deemed permissible, whereas in the third cluster, 2016-2019, three take on the value of permissible. Thus, we now have a picture of how prevalent the publication is of opinion pieces authored by retired military officers in major newspapers which have failed to uphold the norm of non-partisanship and apolitical position.

\singlespace

```{r second level analysis, include = FALSE}

# Present Results of Result and Justification
secondlayer_obs <-adjdataset %>% arrange(PubYr) %>% 
  filter(MilPSoc==1 | Dom_P==1 | Fig_Nme==1 )  %>% 
  filter(Author != "UNK") %>% filter(Author != "Murtha, John") %>% 
  select (Author, Title, Source, PubYr, Result, Justification) %>% 
  view()
```

```{r table3, include=TRUE, echo=FALSE}
secondlayer_obs %>%
  kable(format="latex", caption="Concern for Op-Ed Pieces Authored By Retired Military Officers in US Newspapers, 1979-2020", longtable=T) %>%
  column_spec(1, width="9.5em") %>% 
  column_spec(2, width="15em") %>% 
  column_spec(3:4, width="2.5em") %>%
  column_spec(5, width="5em") %>%
  column_spec(6, width="6em") %>%
  kable_styling(latex_options="striped", font_size=10) %>% 
  row_spec(0, bold=T)
```
\doublespace

Figure Two displays the number of opinion pieces authored by retired military officers which failed to uphold the strict standards of professionalism, as espoused by Retired Generals MacArthur and Dempsey, within each applicable year. The results reveal several conclusions. The first is that it clearly shows, in a broad sense, that from 1979 through the first half of 2020, opinion pieces penned at the hand of retired military officers which violate the concepts of professionalism in questionable or inappropriate ways are the exception, not the norm. Indeed, the entire count of questionable or inappropriate publications after conducting the second level of analysis drop to just 30 of 334 total pieces, or 8.9% of the entire sample. Thus, ample support exists to confirm the first hypothesis, that the vast majority of opinion pieces authored by retired military officers will be apolitical and non-partisan in nature.

```{r fig2, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
datatograph <- data.frame(
  year=c("1998","2000","2003", "2004", "2006", "2007", "2016", "2017", "2018", "2019", "2020"), count=c(2, 1, 1, 3, 3, 1, 5, 4, 4, 5, 1)
)
ggplot(datatograph, aes(x =year, y =count)) + geom_bar(stat = "identity", width=.2) +
  labs(
    x = "Year",
    y = "Count",
    color = NULL,
    title = "Figure 2: # of Op-Eds Written By Military Retirees",
    subtitle = "Failed to Uphold Professional Standards"
  )
```

The second hypothesis, which inferred that because overall circulation among major US newspapers has been declining over the past several decades, the number of opinion pieces penned by retired military officers in these outlets would likewise decline, does not appear to be represented from the data. In fact, we find evidence for the opposite trend: retired military officers are penning opinion pieces at an increasing rate, even as overall circulation among the major newspapers are down in terms of overall circulation (even after accounting for digital offerings) over the same time period.

The third hypothesis, that the number of published opinion pieces which have strained the concept of professionalism has increased, even as political polarization has worsened in the past several decades, is also confirmed, as evidenced by both by Table 3 and Figure 2. It is interesting to note that according to Figure 2, the first publishing year in which an opinion piece penned by a retired officer failed to uphold professional standards occurred in 1998, nearly twenty years into (or almost half) of the entire sample. Why is this the case? Retired military officers in the years preceding 1998 had plenty of reasons to go to the press and air disagreements with political leaders, or to address important topics that may have pushed the limits of professionalism as has been defined in this paper. In fact, many of these officers would have been veterans of the Vietnam and Korean Wars, and perhaps, like Maxwell Taylor, who is in our sample, veterans perhaps of World War Two. At this point, though it is only speculation, one can speculate that the factors of military credibility and political polarization, however imprecise and hazy at this point, may have played a role in keeping the majority of an earlier generation's retired officers out of the popular press.

## Increasing Diversity of Topics Addressed By Military Officers

It is also clear from the data presented in  Table 3 that several publications are deemed questionable or inappropriate from the perspective of professionalism in that they address topics that are potentially outside of the realm of proper military expertise to address. In his _Washington Post_ editorial, "Home Should Not Be a War Zone," Retired Army General Stanley McChrystal addresses the tragic nature of mass shootings in America. He states in the article, "...as a combat veteran and proud American, I believe we need a national response to the gun violence that threatens so many of our communities" [@mcchrystal_home_2016]. In a later article published in the _Washington Post_ in 2018 entitled "Good Riddance," McChrystal shared how his beliefs in Robert E. Lee had evolved over the course of his professional career, and why he now believes it is now time to discard Confederate statues and other similar symbols. Admiral (Retired) Mike Mullen and co-authors Lieutenant General (Retired) Mark Hertling and Admiral (Retired) Robert Natter wrote publications ("The Refugees We Need" and "Cutting Refugee Admissions Will Have Severe Consequences for the US Military") explaining how President Trump's immigration policies on refugees harm military readiness.

While many Americans agree with the stance these officers have taken on what are clearly important issues, one can simultaneously and not unreasonably ask whether it is appropriate for retired military officers to be considered appropriate elite voices on these specific issue areas. Again, the potential issue as it is related to the concept of military professionalism is not that these officers should not have an opinion on these topics, but whether they, as retired military officers, ought to be considered as appropriate mouthpieces by the American public on these issues. It is not unreasonable to insist that policy arenas involving appropriate responses to mass violence, the future of Confederate symbols, and refugee/immigration policy for the nation belong to non-military elites.

```{r fig3, fig.height=4, fig.width=5, fig.align="center", include=TRUE, echo=FALSE}
datatograph <- data.frame(
  decade=c("1979-1989", "1989-1999", "1999-2009", "2009-2019" ), numcount=c(13, 51, 102, 161)
)
ggplot(datatograph, aes(x =decade, y =numcount)) + geom_bar(stat = "identity", width=.2) +
  labs(
    x = "Period",
    y = "Published Opinion Pieces",
    color = NULL,
    title = "Figure 3: Total Number of Op-Eds Written By Military Retirees",
    subtitle = "WSJ, NY Times, Washington Post, USA Today, and LA Times"
  )
```

## Conclusion: Can the Military Remain Professional?

The dynamics of the military institution engaging the American public in a polarized environment remains ripe for future scholarship. In the author's opinion, whether the American military can, and to what degree it will, be able to remain professional as an institution remains to be seen. As political polarization becomes increasingly salient, several dynamics are unfolding that are worth noting, as their full impact has yet to materialize.

The first is that as political polarization worsens, the entire notion of loyalty is being reshaped, redefined, or at the very least, called into question. Consider that in June 2020, following the President's largely unpopular move of displacing protesters from Lafayette Square, a letter published by West Point alumni expressly called out prominent West Point alumni serving in the Trump administration. Though it did not mention individuals by name, the letter was intended and acknowledged by media as directed towards the Secretary of Defense, Mark Esper, and the Secretary of State, Mike Pompeo, who are both graduates of West Point's Class of 1986 [@gaouette_group_nodate]. The charges of the letter were both damning and politically striking: "...We are concerned that fellow graduates serving in senior-level, public positions are failing to uphold their oath of office and their commitment to Duty, Honor, Country"[@line_letter_2020]. The letter's clear implication is that if you are serving in the Trump administration, and if you graduated from the US Military Academy, you have forsaken your oath of office, and by extension, the US Constitution, as well as the values that West Point stands for. Increasing political polarization promises to make these charges of misplaced loyalty more common. Scholarly literature has looked at how the military faces pressure stemming from divided institutional loyalties [@betts_soldiers_1991], but polarization promises to add even more pressure and scrutiny on the loyalty of individual officers and the military as a whole.

A second dynamic, that of maintaining credibility and trust with the American public, remains crucial for the military, but this dynamic will likely be tested in fierce ways should a polarized political climate endure. Even as high credibility with the American public can give the military support, it can also be abused, especially if military officers, to include retirees, use the high level of credibility they currently enjoy to take up and advocate for partisan positions, or worse, to jockey with civilian leaders for power [@brooks_shaping_2008-1]. Unfortunately, as this paper illustrates, there is evidence that mild aspects of this are occurring, but the extent to which it is, and its resulting impacts, remains an area for future research. Credibility is not something that should be shunned. The military as an institution, to include military retirees, holds relevant and military expertise that should be afforded to the American public at large. And yet, just as retired officers were skilled in the management of violence when they wore the uniform every day, retirement in a polarized era means that they must so also be managers of American trust. The promising news is, as this paper makes clear, the great proportion of retirees who are speaking out in various settings know and internalize this. Nonetheless, this will become more difficult to do in a polarized environment.

A third, and perhaps the most important dynamic of polarization as it concerns the military, is that the line between what is considered partisan or political and what is considered morally right and wrong is becoming seemingly blurred, if not altogether disappearing. Polarized sides tend to believe that each is acting "rightly" and that the other side is not just mistaken, but deeply wrong in a moral sense.

These three factors are difficult in a purely domestic political environment, and one does not have to imagine how the potential snares of a polarized political environment and military credibility stewarded poorly could impede the military's ability to do it's main task, fighting and winning this nation's wars. Indeed, the impact of these characteristics on our nation's military effectiveness must also be considered in further research. Regardless, the importance of individual moral courage will continue to grow in value as an attribute of military leaders. Its importance as a trait will likely continue to grow in the years ahead.

\singlespace
\pagebreak