### 2020-10-01 code/prospectusv1.Rmd
commit

---
title: "Loyalty's Limits: The Impact of Postmodernism and Political Polarization on Military Professionalism  (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation seeks to explore the impact of increasing political polarization on the professional identity of military officers and the military institution. I develop a theory that argues that the nature of rising political polarization fundamentally challenges the conception of loyalty among members of the military profession, and that such a strain is problematic for an institution such as the military that holds the virtue of loyalty in high regard, and by statutes possesses multiple, simultaneous loyalties: to the Constitution of the United States, to the President qua Commander-in-Chief, to Congress, to the American People, to one's fellow service members, and to one's self. Balancing these competing loyalties becomes more difficult as political polarization worsens, I argue, because of the fundamental and core values - not mere policy differences - that are at the root of many contemporary polarized political battles. Employing a mixed-methods approach that incorporates quantitative analysis and several historical case studies, this dissertation argues that the challenges to conceptions of loyalty among members of the military institution brought about by political polarization have the potential to impact normative concepts of military professionalism, which could be problematic not only for the character and identity of the military as an institution, but for American national security at large._   
geometry: margin = 1.15in              # margin widths
fontsize: 12pt                         # text size
bibliography: researchpaper-ps904.bib # bibliography file
biblio-style: authoryear               # citation style
urlcolor: blue                         # links to external URLs
citecolor: black                       # links to bibliography
linkcolor: magenta                     # internal figures, footnotes, etc.
indent: true                          # paragraph format
                                       # true for indented paragraphs

header-includes:
    - \usepackage{setspace}
    - \usepackage{multirow}
    - \usepackage{indentfirst}

output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    number_sections: false
    toc: false
    fig_caption: true
    citation_package: biblatex
    includes: 
      in_header: 
        - latex-preamble.tex
  bookdown::htmldocument2: default
  bookdown::word_document2:
      toc: false
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
library("broom")
library("dplyr")
library("reshape2")
```
\newpage

## Introduction and Research Question

In late May and early June 2020, racial tension and concerns about police brutality gripped the American public following the death of George Floyd, an unarmed black man, by police in Minneapolis. In many cities and communities across the nation, peaceful protests were followed in some cases by periods of rioting and looting. On June 1, after giving remarks at the Rose Garden, President Donald Trump walked to St. John's Church after security forces visibly displaced throngs of protesters in neighboring Lafayette Square. Flanked by members of his cabinet, to include the Secretary of Defense, and with the Chairman of the Joint Chiefs of Staff not far behind, the President held up a Bible at the church, stared into the cameras of the media, and, several minutes later, returned to the White House.

In the days and weeks that followed, a full-blown civil-military relations crisis ensued  [@feaver_military_2020]. The Secretary of Defense, Mark Esper, publicly disagreed with a threat made by President Trump to invoke the Insurrection Act of 1807 to use active-duty forces to quell domestic riots [@esper_secretary_2020], and General Mark Milley, the Chairman of the Joint Chiefs of Staff, told senior military officers graduating from the National Defense University that he "should not have been there (accompanying the President on the walk to St. John's Church)", as his presence "created a perception of the military involved in domestic politics" [@milley_official_2020].

Several retired military officers also weighed in, not only on the Trump administration's handling of specific events near the White House, but on a host of broader topics that dominated the popular dialogue in America after the death of George Floyd, to include the state of American racial relations. Upon paying close attention to what these retired military officers said, though, one can notice that there was in fact fairly substantial variation in both the substance and tone of their remarks. While all of them expressed concerns about the President potentially ordering active-duty military forces to confront what many Americans believed were peaceful protesters, and thus, violating Constitutional protections afforded by First Amendment [@brooks_dismay_2020; @dempsey_former_2020], some officers went much further in their criticisms. Indeed, some of these retired military officers accused the President of failing to possess the overall fitness to lead [@goldberg_james_2020], promote harmonious racial policies  [@allen_moment_2020], and demonstrate adequate leadership abilities [@mullen_i_2020]. In the months following George Floyd's death, moreover, several retired and currently-serving military officers have further made public arguments on a variety of topics that are directly related to President Donald Trump, including race in America [@askew_anti-racist_2020] and voter fraud as it relates to the upcoming Presidential Election [@nagl__2020].

While America has witnessed numerous incidents in its history in which tension between civilian and military elites has been present, the events of the Summer of 2020 between the Trump administration and some - though certainly not all - retired  military officers raises important questions, particularly I argue, because the context did not involve combat overseas.  These questions include: why did several retired military elites speak out so strongly against President Trump? Why did some of these elites confine their criticism to concerns over the Constitution, while others sounded more like representatives of the Democratic Party? Are these events that occurred during the Summer of 2020, and the military response generated in response to these events, an isolated chapter in American history, or is there something more disturbing underneath the surface regarding contemporary American civil-military relations?

This dissertation seeks to explore the impacts of rising domestic political polarization on the US military as an institution and profession. The primary argument espoused in this dissertation is that certain features of the polarized environment impact how military officers view and exercise the concept of loyalty, a claim that I argue holds significant ramifications for US national security. Using a mixed-methods approach, I develop a theory and then test it using a mixture of quantitative analysis and several qualitative case studies.  I conclude with discussing the implications for such an argument.

# Framing Within the Scholarly Literature

## Political Polarization

The question of if and how political polarization drives changes in the professional identity of the military spans several related literatures, most notable of which are American Politics and International Relations. Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature. Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics. I will frame my research question within each of these fields briefly, and to define important concepts as required, before addressing why this question matters and should be studied in depth.

Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012]. Despite the wide-ranging scholarship exploring the many facets of political polarization, virtually all scholars agree that contemporary levels of political polarization in the United States are historically abnormal and pose myriad problems.

One of the major concerns regarding political polarization is its on impact on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world. Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017]. Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions will yield findings that may help policymakers and academics help solve the problems Schultz identifies.

## Civil Military Relations

Civil-military relations comprises an immense area of scholarship and interest, and deals with a correspondingly large number of questions and debates that have occupied the minds of political scientists, theorists, historians, ethicists, and sociologists for hundreds if not thousands of years. At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or how to ensure that the military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999]. Feaver describes his "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p. 214].

Civil-military relations involves a wide angle of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority, and a host of other massive and consequential issues impacting how militaries perform in war, their relationship to the societies they serve, and other important questions. In the context of the US, which has never experienced or come close to a coup (thankfully), the study of civil-military relations often focuses on the relationship between military and government elites, or what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p. 380]. Indeed, it is at this level that vital decisions are made, including to go to war, how to prosecute war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be. Each of these questions, and many, many more, involve questions regarding the proper role of the military and its leaders relative to both society and, perhaps more important, the leaders of government. A number of scholars have examined this question on normative grounds, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for many scholarly works on civil-military relations. Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's concept of normative civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes the concept of what Huntington called "military professionalism" [@huntington_soldier_1957].

In _The Soldier and the State_, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States. The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3]. The societal imperative, Huntington claimed, was a further function of two distinct features of the United States. The first was an "ideological constant," which referred to America's political ideology of liberalism [@huntington_soldier_1957, 143-162]. Huntington argued that liberalism as a political ideology stressed "individualism" [@huntington_soldier_1957, 90], and that American society was, since its founding, suspicious of standing military organizations and professional military forces. The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143]. The tension raised by these two imperatives - one functional stemming from external threat, the other societal stemming from the character of American political ideology and structure - needed to be solved, claimed Huntington. Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society. Both goals - military security and civilian control - had to be obtained.

Figure 1, below, shows a visual diagram of Huntington's logic in a political science framework [@feaver_civil-military_1996]. As Peter Feaver points out in a 1996 article, Huntington did not use such language when developing and explaining his concept. According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military. Huntington viewed the key independent variable that impacts and enables these goals while also resolving the tension generated by the opposing functional and societal imperatives, as described above, the adoption of a model of "objective civilian control" of civil-military relations, whereby political and military leaders "maximize military professionalism," which includes granting the military a high degree of professional autonomy [@huntington_soldier_1957, 80-85].

```{r figure one, include=FALSE, echo=FALSE}
huntington_dag <- dagify(
    MSCS ~ CC,
    CC ~ FI + SI,
    SI~I + S,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MSCS", 1, .5,
      "CC", .7, .5,
      "FI", .5, .6,
      "SI", .5, .4,
      "I", .3, .3,
      "S",.7, .3,
    ))

huntington_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

huntington_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MSCS"="Military Security and\n Civilian Supremacy", 
                     "CC"="Patterns of\n Civilian Control\n 'Objective Control'", 
                     "FI"="Functional\n Imperative", 
                     "SI"="Societal\n Imperative",
                     "I"="Liberal Ideology",
                     "S"="Constitutional Structure")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory\n Of Civilan Control of the Military", subtitle="As espoused in The Soldier and The State, 1957")

```

Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83]. In fact, Huntington admits that the US Constitution actually envisioned this "subjective" scheme of civilian control of the military [@huntington_soldier_1957, 163-168]. What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats. It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and battled during the 19th and 20th centuries during World War One, World War Two, and now the Cold War, that the military profession had risen as a separate and distinct body that could focus on the demands of military service. Such a separation and distinctness, claims Huntington, would need to be embraced by civilians and the militarym going forward into the future. In so doing, claims Huntington, is the only way that both military security and civilian control of the military - two often-competing imperatives - will be safeguarded.

Numerous critiques surrounding the Huntingtonian ideal of _military professionalism_ have been levied over the years, as Huntington did not attempt to precisely define it. Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres is fundamentally flawed. Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], charging that war by its very nature requires the highest integration - not separation - of military and political activities. In essence, these critiques argue that the strict separation of the responsibilities between the military and civilians leads to poor strategic outcomes [@rapp_civil-military_2015; @brooks_paradoxes_2020].  Moreover, these criticisms can and do point largely to American failures in Iraq and Afghanistan to secure, at least as of yet, a more-lasting peace - failures that are ultimately, according to one scholar and former general, moral in nature, and shared between both civilians and the military [@dubik_just_2016].

As an extension of the critiques of Huntington, there has also been renewed focus and attention among civil-military relations scholars on numerous professional normative standards of the military. One of the most important -  yet difficult - debates is an attempt by scholars to define where appropriate political behavior ends and partisan activity begins [@owens_military_2015; @golby_beyond_2015]. In the Huntingtonian framework, this is essentially the opposite side of the debate over the same coin in attempting to define for civilians what constitutes meddling in military affairs. One of the reasons this debate has become even more difficult, I believe, is because as political polarization increases, the lines between partisanship and politics, as well as politics and everything not politics, have blurred, as so much of the literature on American politics and polarization affirms.

## Military Professions and Organizational Loyalty

Loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative sense and in the context regime exclusion and a host of related phenomena, including military coups [need cite], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998]. This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader. Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that perhaps are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the military. Because appropriate to explore partisan or ideological (instead of ethnic) identification stemming from However, this literature often assumes a definition of loyalty and rarely develops how the reader should think of loyalty in the first place. For this fundamental starting point of defining loyalty, the fields of ethics, management, and sociology are helpful.

Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019]. Though the differences between the two approaches are important, both assert that loyalty is something felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, p. 109]. Oldenquist, a philosopher, describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it [the object of one's loyalty] prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, p. 175, brackets mine].

Loyalty further generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p. 175] to the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992]. Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009]. For organizations, and by extension, professions, the concept of loyalty is also very important. Conner et al. state, for example, that "having loyal members is vital to perpetuating the group" [@connor_military_2019].

We also know that the military services in the United States place a premium on loyalty as a trait and concept. Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty constitutes a significant part of the Army profession [@us_department_of_the_army_adp_2019]. This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army Profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019]. There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, or "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p. 1-9.].  The US Army states in its doctrinal manual on Army Leadership and the Profession, ADP 6-22, that the Army Values "embody the practical application of the Army Ethic. They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p. 1-12]. The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [@us_department_of_the_army_adp_2019]. The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception. In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918]. Even Huntington spent a great deal talking about the importance of loyalty as a trait to the very essence of the military profession. He recognized that in order for civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].

Loyalty is further important in a military setting in a broader and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as an apolitical and non-partisan institutions with multiple, simultaneous loyalties. The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, deliberately contrives the military as an institution possessing multiple loyalties. One of the most basic forms of these multiple loyalties is captured in the fact that the Constitution empowers the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also gives Congress an array of tremendous powers and responsibility pertaining to the Armed Forces @[noauthor_us_1787-1]. Indeed, the tensions that inherently exist in American's constitutional framework are explored in a robust literature that examines the multiple ways in which the Congress, the President, and the military interact to implement, but also to influence, American foreign policy [for instance, see @howell_while_2007].

In the context of the US military as an institution, loyalty has rarely been used as a conceptual construct in the political science literature, likely and thankfully because the US military has not yet considered, or been involved, in a coup. Nonetheless, I contend that there is a rich opportunity to study strategic loyalty as a concept, especially in a contemporary context in the United States, where political polarization has increased extensively.

# Where My Puzzle Fits and Why It Matters

One area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity of the military, both as an American institution and as a profession. Exploring this relationship will, I believe, help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate. On one hand, one would expect that it might, in the sense that the American military is, after all, a microcosm of American society. Therefore, one could expect that if society truly is becoming polarized, then at least to some degree, so is the military, or at least those who volunteer for service and at the point in time that they join. On the other hand, one might reasonably expect that the unique purpose of the military, combined with its long-standing professional ethic, might insulate the military institution from some of the potentially negative impacts of polarized politics. These differing expectations need to be more clearly articulated and explored.

I am specifically interested in the question of if and how political polarization impacts the concept of organizational loyalty in the military. This is an interesting research question because, as noted above, one of the most fundamental normative hallmarks of the military's professional nature is that it ought to be apolitical and non-partisan, and one of the features of militaries in general is their tendency to foster a concept of loyalty among those who serve. Yet a puzzle emerges: within an institution that is normatively apolitical and non-partisan, can it remain so in an environment marked by increasingly polarized domestic politics, where partisan loyalties are a salient feature of the political landscape [need cite here for partisan loyalties].

My intended dissertation will help answer several related questions, including: if personal and partisan loyalties are changing as a result of political polarization [need cite], is this also happening within the military institution, and why? Are partisan/ideological loyalties displacing other loyalty relationships within the military? If so, are these changes in organizational loyalty having an impact on the military's professional identity and character? Through what mechanisms do these changes operate? What sorts of changes are they driving, if any, in the military institution? Is this problematic for the military profession, and why?

The answers to these questions, which I intend to include as part of my dissertation, will help inform scholars and policymakers alike on number of important issues. First, it will help illuminate in a broad sense an important aspect of how changes in political polarization impact American institutions, and in particular, the military. Second, work on this topic will attempt to explore how a fundamental concept and value within the military - loyalty - is both impacted as a result of political polarization but also how loyalty changes may be driving changes in the military's professional character, identity, and nature. In doing so, it will help enrich a scholarly understanding of what loyalty is to begin with, and how it features prominently in the context of the American military.

To be clear, scholars have long studied how domestic politics impact the military. And as scholars have pointed out, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996]. And as American history proves, there have been countless instances in which civilian and military leaders disagreed. Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan. Beyond these episodes in which President and wartime commander locked horns, consider several other instances of "civilians" and "military" clashing. One episode is certainly the way in which the Vietnam war escalated in ways that, at least according to H.R. McMaster, military generals at the time knew were destined to fail [@mcmaster_dereliction_1998]. A more recent example, and one much studied and commented upon by civil-military relations scholars, is the so-called "Revolt of the Generals" that occurred following the early years of the invasion of Iraq, which involved several recently-retired generals and admirals sharply criticizing Secretary of Defense Rumsfeld's leadership dynamics [@cook_revolt_2008].

My personal motivation in choosing the intended topic, however, stems from what I believe are changes in the type of civil-military disagreement that have surfaced in recent years. For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, way outside of appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation] More interestingly, these conflicts are occurring not in the context of major war, but the opposite: a period in which the levels of US Troops deployed in active combat zones is less than it has been in recent years. While I believe that such developments are catching the attention of civil-military relations scholars, most of the work done on this topic is, in my opinion, impressionistic [@golby_jim_keep_2020] as opposed to systematic. In other words, civil-military relations scholars have identified that polarized partisan politics serve as a risk for the military, but none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and how.

Thankfully, a growing number of civil-military relations scholars have looked at various aspects of this question - the impact of the political polarization on civil-military relations - in recent years. But the focus of these works has been, I believe, not on the impact of polarization on the military itself but on other important but peripheral components of the military's relationship with American society. One recent dissertation completed by Michael Robinson at Stanford University in 2018, for instance, found that partisan affiliation significantly affects public credibility and trust in the military, perhaps resulting in or as a consequence of media "echo chambers" which may incentivize military leaders to offer opinions partisan opinions without fear of sanction [@robinson_danger_2018]. Robinson's work builds off of previous work by Heidi Urben in which she found that active military officers believe less strongly that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013], a norm that has existed for decades or more. Together, Robinson and Urben join other scholars in providing a wealth of information regarding public opinion and the military, and in particular, how aspects of political polarization shape partisan differences regarding topics such as the use of force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].

# The Analytical Framework - Initial Argument

## How Polarization Impacts Loyalty as a Concept

In this section of the prospectus, I will lay out an initial argument, as well as accompanying hypotheses, which I intend to explore and test in my dissertation. Figure 2, below, shows an initial conception of the linkages between several of the variables discussed in the previous section. From left to right, this conceptual model suggests that worsening political polarization is driving changes in both the institutional credibility of the military (how society views the military) and changes in the conceptions of loyalty and identity (within the military). These two independent variables are, I suggest, driving subsequently driving changes in the concept of military professionalism, the main dependent variable of analysis in my dissertation. Following Huntington's model, I also suggest that changes in military professionalism can impact a variety of outcomes, by which I mean the very same outcomes that other scholars do and have already explored, such as effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] and achieving strategic effectivness [@brooks_creating_2007; @brooks_shaping_2008]. While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation, as well as the least-explored linkage in this model, is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism. For emphasis and clarity, I do not think that the other link - that between political polarization and institutional credibility - is unimportant. Rather, this link seems to be the focus of more-recent scholarship [@robinson_michael_danger_2018].

```{r figure two, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    MO ~ MP,
    MP ~ LI + IC,
    IC~PP,
    LI~PP,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MO", 1, .5,
      "MP", .7, .5,
      "IC", .3, .3,
      "LI", .3, .7,
      "PP", .1,.5,
    ))

problem_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MO"="Changes in\n Military\n Outcomes", 
                     "MP"="Military\n Professionalism", 
                     "PP"="Increasing\n Political\n Polarization", 
                     "IC"="Institutional\n Credibility",
                     "LI"="Concepts of\n Loyalty and Identity")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="The Impact of Political Polarization on Military Professionalism")

```

A primary contention from the system visualized above is that as political polarization in the domestic sphere increases, the military faces increased challenges in how both individual military leaders as well as the institution conceptualizes, adopts, and exercises loyalty, and by extension, professional identity. This expectation stems from several features that scholars of American politics have raised regarding the current politically polarized climate within the United States. First, scholars have noted that political polarization raises the number and scope of issues that are subsumed by partisan debate and disagreement (a phenomenon referred to among scholars of American Politics as "conflict extension" [@layman_party_2002]). A second feature of contemporary political polarization is the prominence of  negative feelings held by one partisan group towards another, a phenomenon referred to as "negative partisanship" [@abramowitz_great_2018]. A third feature is an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018]. I contend that these features impact individuals serving in the military, including military leaders, as they do members of the domestic public writ large, and that these impacts are problematic for members of an institution that normatively exists as an apolitical and non-partisan actor. Table 1, below, lays out how these three prominent features of domestic political polarilzation may impact the non-partisan and apolitical nature of the normative military professional and institution.

\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{How Features of Contemporary Political Polarization Impact Normative Military Professionalism} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Professionalism}  \\ \hline
    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number/scope of issues. & Military Professionals and the Institution find it increasingly difficult to "stay out" of partisan politics as the territory of what is considered partisan politics grows. \\ \hline
    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain arise against those who hold to opposing political views. There is a belief that those on the other side of the political aisle are not just mistaken over policy, but greatly wrong and flawed as individuals. & Even among military leaders, there is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
    \textbf{Individuals believe that "core" issues are increasingly at stake} & The salience of one's identity and core values is raised as partisans believe essential issues are at stake in the political arena. Associated with this is a belief that not just policy preference, but moral right and wrong, are at stake. & Military professionals, who hold multiple institutional loyalties, face an increasingly difficult task of balancing these loyalties, especially as the objects of these loyalties are increasingly associated with or attached to notions of moral right and wrong. \\ \hline
\end{tabular}
\label{table:comparison} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

The first way in which the apolitical and non-partisan ethos of the military is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism. This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is simply less politically-neutral territory available. Thus, an actor who wishes to remain non-partisan, such as the military, may be increasingly less able to do so, as a result of this feature.

A second challenge military leaders and the institution faces in a polarized climate is from the "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape. As citizens hold increasingly strong and negative feelings towards those who side differently on a host of issues, there may be a tendency for members of the military institution to harbor similar feelings. Even if impact of this feature on the military and its leaders, due to training and professional development that emphasizes the importance of upholding apolitical and non-partisan behavioral norms, is somewhat muted, the point is that members of the military institution are likely experiencing the same phenomenon - even if it is to a lesser degree - as the country's citizens.

A third, and I contend the most impactful, challenge that normative military professionalism faces stems from the belief by individuals that "core" issues are at stake in the current polarized political climate. "Core" issues are difficult to define precisely [need cite], but the main point is that these issues often involve perceptions of moral right and wrong, and not just simple policy preferences. Moreover, some of the most polarizing issues are often associated with issues related to an individual's identity, such as race, gender, immigration [need cite]. The rise of "identity politics" in America as political polarization has worsened illustrates this point [need cite].

I argue that this third challenge is especially relevant for the concept of loyalty. As an increasing number of what individuals believe to be "core" issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects. I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders. Not only is the de-facto setting in which military leaders operate marked by multiple and often conflicting loyalties due to the structure of and statutes set forth in the US Constitution, but also, the potential consequences are higher for military leaders than average citizens should they fail to discharge the duties that correspond to their roles as professional officers. These consequences include failure to achieve tangible national security goals, such as loss on the battlefield, or less tangible but still important outcomes such as maintaining trust with the American public.

Moreover, as important as these consequences are, they do not speak to the potentially high personal and individualistic consequences military leaders may face if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best. Military leaders are still humans who have to live with themselves at the end of the day. Therefore, I argue that if polarization continues to erode boundaries of what is considered moral by individuals to begin with, there is an increased likelihood of military leaders having to choose between conflicting loyalties. What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically? Can such an outcome exist, and if it can, is there empirical proof that such a trend is occurring? If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?

## How Polarization Interacts with Institutional Credibility

The previous section described how increasing political polarization presents significant challenges to aspects of the normative concept of military professionalism. But whether these challenges enter the public domain, or remain private challenges that specific individual actors contend with on a personal level only, matters a great deal because it "changes the character...from a matter of principle to a political matter" [@dubik_principled_2014] In fact, in much of the field of civil-military relations, this dynamic of private disagreement or discontent potentially entering the public realm is seen in the scholarly literature on senior officer resignation. This debate examines various aspects of resignation, including whether there can ever be conditions under which resignation is justified or appropriate, with scholars taking various sides on the issue, both for [@dubik_principled_2014; @snider_strategic_2014] and against [@feaver_should_2014; @golby_beyond_2015]. I contend that polarization and credibility of the military interact, however, in at least two important ways. First, as loyalties are challenged as a result of increasing political polarization, individual military leaders may not necessarily remain quiet despite the existence of a strong professional norm of doing so because the objects of loyalty may demand that their adherents publicly speak out. This of course, depends specifically on what the object of loyalty is and what it demands from those who hold closely to it. The second way in which polarization and credibility may interact arises more directly from what can be termed a credibility effect, and has more to do with the place of prominence (or lack thereof) that the military or individual leaders hold in the perception of American society. [need to blend this in with recent scholarship and where gaps are - bring in Urben, Golby, Dropp and Feaver, etc., and Robinson.  Also need to figure out where to put my hypotheses]

I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed. From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance). From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high. By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.

When military credibility is low, the theory contends the opposite will occur. Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists. At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to. In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public. Such interaction can take a variety of forms. When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.

Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 2, below). The first is when political polarization and military credibility are both low. When these conditions hold, there is no anticipated drop in military professionalism by the military institution nor its individual elites. With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues. 
\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
    \textbf{Low Political Polarization} & Negligible anticipated impacts. Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts. There is a demand for military voice, but little substantive matter stemming from US domestic politics that can cause crucial disagreement. Little incentive to deviate from traditional concepts of loyalty and duty. \\ \hline
    \textbf{High Political Polarization} & Possible Impacts. Polarization raises the stakes of domestic politics, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism. Likely impacts. Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between individual military elites and civilian leaders on a number of crucial topics.\\ \hline
\end{tabular}
\label{table:comparison} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

The second possible set of conditions is when political polarization is low and institutional credibility is high. When these set of conditions exist, the military has a platform to interact with the public, as the public highly respects the military institution. The military enjoys this high level of trust and does not want to lose it. Further, because the degree of political polarization is low, there is little incentive for the military as an institution or military elites to use the trust they enjoy with the American public to adopt or denounce an issue associated with any partisan entity. The high institutional credibility increases the chance of the military interacting with the public in an unprofessional way, but the low political polarization removes the incentives of doing so.

The third type occurs when political polarization is high, but institutional credibility of the military is low. Under these conditions, the military does not command the respect of the American public like it does when credibility of the institution is high, and thus, does not possess a considerable outlet to interact with the American public. However, the incentives of loyalty and duty still loom large in the minds of some military elites as a result of the high degree of political polarization, and thus, a dip in professionalism is still possible.

The final set of conditions exists when political polarization and institutional credibility are high. It is in this world that I contend the American military finds itself in now in 2020. Under these conditions, the military faces dual dangers. High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very traits it is most adhered to - loyalty and duty. Indeed, this set of conditions comprise a very dangerous world for the military institution and the military professional.

# The Analytical Framework - Hypotheses

In order to test my argument, I will have to examine and test several hypotheses, which are inter-related and linked. They include the following:

_H1: In highly polarizing times, there will be an increase in the frequency of military officers breaking normative civil-military relationship behaviors, such as criticizing the President, members of Congress, or those who hold to different political views on particularly-polarizing issues outside of the conduct of war._

_H2: These breaks in normative civil-military behaviors will occur because political polarized periods feature events or situations which test or challenge the military's ability to juggle and balance multiple simultaneous loyalties: to the President as Commander in Chief, to Congress in its oversight role of the military, to the underlying ideologies or ideas inherent in issues that prevail during polarized times, to the normative professional ideals of the military, and to the individual's bonds of loyalty to his or fellow soldiers and the mission._

_H3: Because of the normative strength and prominence of place that the ideal and concept of loyalty has within the profession of arms, military leaders are more sensitive than regular civilians to appeals of loyalty._

_H4: Following from the above hypothesis, polarized periods will feature events in which governmental actors appeal to the military's sense of loyalty in order to persuade them to adopt a stance on a particular issue area._

_H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution changes. If changes occur to the military institution, they are bound to occur within two primary spheres: the military's professional identity, and the military's professional coherence._

# Plan of the Dissertation and the Data Generating Process

To complete my dissertation, I plan on incorporating a mixed-methods approach. In broad conceptual terms, I plan to propose a theory and then test the theory using a combination of quantitative analyses and qualitative case studies. The one sentence thesis of my dissertation is that political polarization significantly impacts how the military thinks about and exercises loyalty, a chief military trait, and that these impacts can and often do result in changes to the military's professional identity and character. I plan to execute a book style project comprised of at least seven chapters.

In the first chapter, I will frame my research question, lay out my research agenda, and review the relevant literature for my project. This will involve synthesizing several of the key aspects and debates within the field of civil-military relations, much in the same way I have done in the first part of this dissertation prospectus. I will focus my attention on introducing and summarizing the concepts of military professionalism, organizational loyalty, and political polarization, and spend time orienting the reader to the debate within scholarly literature regarding a Huntingtonian conception of military professionalism. I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020]. The main point of chapter 1 is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: it may insulate the military from the most pernicious effects of polarizing politics.

In the second chapter, I will introduce, develop, and explain my theory that links increasing political polarization to conceptions of loyalty and military professionalism (see Figure 2). I will argue that specific features of the polarized environment strain and challenge both individual and institutional conceptions of loyalty, and by extension, normative aspects of military professionalism (Table 1). I will further argue that rising political polarization can interact with changes in the military's institutional credibility to increase the likelihood that the public becomes aware of changes in the military's professional conduct. Using Table 2 as a framework, I will place epochs of historical American civil-military relations into one of four distinct quadrants, and signal that these quadrants will comprise the qualitative case studies that I later examine in my dissertation.

In the third chapter, I will argue that today, in 2020, the American public is witnessing a decline in professional military conduct among retired military officers. I plan to demonstrate this empirically by drawing upon an original dataset consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times. I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more importantly, that there has been a decline in the normative professional ideals espoused and argued for in these op-eds. Critical to this chapter is acknowledgement that coding for what is professional versus what is not is subjective to a large extent, and thus, the need for me to truly develop clear working definitions of the concept of military professionalism in Chapter 1. However, I will show that whether one adopts a strict Huntingtonian version of military professionalism, or a less strict version that allows for more encroachment by the military institution into the political sphere, by both standards the retired military officer corps is not living up to the ideals of normative military professionalism. The question, then, becomes why aren't they? I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the drop in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty. Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vangaurds, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p. 1-9] of arms.

In the fourth chapter, I want to show that the institutional challenges to loyalty have been present as political polarization has worsened. To do this, I plan to examine, via textual analysis, the records of Congressional Committee hearings in which military officers, active and retired, have played a part, and detect if and how patterns of questioning and providing answers have changed over time. I am especially interested in the way that individuals have invoked language associated with the concept of loyalty in asking and answering questions. An example of this might be analyzing the ways in which terms such as "moral obligation" or similar language is used in these hearings. I hope to detect patterns that will help support my overall argument. I have yet to flesh this particular chapter out some more.

In Chapter 5, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers. In this survey experiment, I plan to test whether members of the military (which I proxy by veterans) genuinely are more sensitive to appeals to loyalty, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age, will present the respondent with 6-8 hypothetical situations and a corresponding argument made by a military elite. Members of the treatment group will receive an appeal to loyalty as part of the elite's argument as to why a certain course of action should be followed regarding the hypothetical situation presented.

I plan to structure the survey such that the scenario and corresponding course of action contain references to issues that are highly correlated, in some cases, to partisan / ideological affiliation.   An example of a question might be as follows:
\singlespace
* Situation: The President of the United States, a [Republican, Democrat] recently commented that he is seriously contemplating ordering active duty military Troops to assist local police and National Guard units in several US cities in maintaining law and order as a result of initially-peaceful protests which turned violent in some, but not all, cases. These protests have been going on for more than two weeks and are largely, but not exclusively, affiliated with [the Black Lives Matter Organization, a pro-2nd Amendment organization].

* Elite Voice: A retired US Army 4 star general, who has extensive combat experience in Panama, Iraq, and Afghanistan, recently penned the following op-ed piece in a prominent newspaper:

* “The military must not be ordered to put down riots and protests, even if a small number of them have, unfortunately, turned violent. Even if the President is legally able to issue such an order, the military’s fundamental purpose is to defend the nation from external threats, and not to endanger the Constitutional rights of everyday Americans. [Treatment group, add ]: Should the President order the active military to deploy to cities where these protests have turned violent, military officers should resist such an order and if necessary, resign their commissions. They owe a higher loyalty to the Constitution, and the President’s order would clearly violate the oath of office these officers have sworn to uphold.”

* Do you find this argument compelling (7 point scale of (1) not at all - (7) absolutely, very compelling).

\doublespace
Such a survey will enable me to measure two things. First, it will allow me to measure whether appeals to loyalty are truly more effective than standard appeals to action. This result can be measured by comparing the answers between the treatment and control groups. Second, it will allow me to measure whether members of the military (or veterans) react differently to these appeals to loyalty than regular civilians. As I have hypothesized in _H3_, I suspect that they do. Of note, I am planning to also pursue a survey among actively-serving military officers, but the lead time for obtaining approval to do this can be very long, perhaps 9-12 months or more. If I am able to obtain approval for this, I will then have three different populations to measure: civilians, veterans, and actively-serving military officers. Of note, if I discover throughout the course of my survey that loyalty does not in fact have the strong explanatory power that I think it does, or if veterans and/or active military officers do not respond to appeals to loyalty any differently than civilians, I think such a finding is also interesting.

In the sixth chapter, I will turn to a series of qualitative case studies to explore and test my theory. Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institional military at large, in line with the theory that I propose in Chapter 2. My initial projection of what these case studies will be and how they will align with varying levels of political polarization and instititional credibility are as follows:

* High Polarization, High Institutional Credibility: civil-military relations in the Andrew Johnson administration after the American Civil War. I will explore how General Ulysses Grant confronted challenges of loyalty to multiple institutions and objects as he dealt with a President who was by many accounts sympathetic to southern interests as the US Army played a prominent role in reconstruction in the south [@hyman_santon_1960].

* High Polarization, Low institutional credibility: civil-military relations in the post-Vietnam period, in the late 1970s.

* Low Polarization, High Institutional Credibility: civil-military relations after World War Two, in the late 1940s and early 1950s.

* Low Polarization, Low Institutional Credibility: TBD.

In the seventh and final chapter, I will conclude by exploring the implications of my argument and by

\singlespace
\pagebreak