### 2020-10-23 code/prospectusv2.Rmd
commit

---
title: "Saviors or Syndicates?: The Rise of Domestic Political Interventionism by Retired Military Leaders (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation explores the changes in how a powerful block of military professionals - retired senior military leaders - intervene in domestic contemporary politics. I advance a theory which argues that changes in political activism by retired senior leaders is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility that military leaders enjoy within the society. To support my argument, I trace changes in levels and types of distinct forms of political activism, including authoring opinion-editorials in major newspapers, and participation in presidential endorsements. I further advance the theory by conducting a survey experiment that tests for differences between civilian and military responses after exposure to situations that evoke one's partisan or ideological identity. I ultimately argue that the increasing political activism of retired military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends that have been active for some time in which both the ideological differences between Americans and the prestige of the military have increased. I conclude by describing the effect of the implications of growing politicla activism by senior military leaders on the core beliefs of the military regarding its very own essence and role in society._

geometry: margin = 1.15in              # margin widths
fontsize: 12pt                         # text size
bibliography: researchpaper-ps904.bib # bibliography file
biblio-style: authoryear               # citation style
urlcolor: blue                         # links to external URLs
citecolor: black                       # links to bibliography
linkcolor: magenta                     # internal figures, footnotes, etc.
indent: true                          # paragraph format
                                       # true for indented paragraphs

header-includes:
    - \usepackage{setspace}
    - \usepackage{multirow}
    - \usepackage{indentfirst}

output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    number_sections: false
    toc: false
    fig_caption: true
    citation_package: biblatex
    includes: 
      in_header: 
        - latex-preamble.tex
  bookdown::htmldocument2: default
  bookdown::word_document2:
      toc: false
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
library("broom")
library("dplyr")
library("reshape2")
```
\newpage

# Introduction and Research Question

Across the world and in almost nearly every country, discussions regarding the participation of the military in politics are nothing new. Indeed, few topics evoke such strong opinions and emotions than whether and to what degree the military, both as a corporate body and as individual military officers, should be political vice remain separate from certain aspects of and activities associated with politics. These normative questions rightly form a significant portion of the study of civil-military relations, and intellectuals, practitioners, and pundits have frequently weighed in on these topics over the years.

Thankfully, within the United States, unlike many other countries across the world and throughout history, there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves. However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether. Indeed, a cursory examination of the relationships between Lincoln and MacClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.

What has changed, however, is the degree to which a very specific subset of the American military institution - retired senior military officials - have participated in the domestic political process post retirement. Indeed, whereas General Omar Bradley, one of the military's five star generals and a hero of World War Two, famously remarked in 1959, "I am convinced that the best service a retired general can perform is to turn in his tongue along with his suit and mothball his opinions," [@times_bradley_1959], retired generals and admirals today appear frequently in the media, endorse Presidential candidates at higher frequency [@griffiths_not_2019], and frequently serve in government after their retirement from military service [@brooks_dont_2016]. To provide one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the use of active duty military forces in American cities, former Chairman of the Joint Chiefs of Staff Admiral (Ret) Michael Mullen was one of at least ten four star retired officers who publicly aired concern about the President's potential use of force. ^[Other officers included General (Ret) James Mattis, General (Ret) Mark Kelly, Admiral (Ret) William McRaven, General (Ret) Martin Dempsey, General (Ret) Colin Powell, General (Ret) Michael Hayden, General (Ret) David Petraeus, General (Ret) John Allen, and General (Ret) Vincent Brooks.]. During an interview in the days which followed, when asked by NPR reporter why he spoke out, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].

Mullen's words are incredible from a number of important perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019]. Few, if any, scholars have attempted to systematically examine and explain the rise of political activism by retired military officers over time. Such an effort would better illumine the understanding of scholars, as well as the American military and society, of why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.  More important, a systematic study of this trend would contribute to a much-needed discussion regarding the role of the military institution and its relationship to the government and American society in a contentious political age.

The central argument made in this dissertation is that increasing political activism at the hands of retired military officers is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and rising prestige and credibility of military institutions afforded by American society. The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient servants and towards "guardians of democracy" who pick and choose when and how to intervene in American politics. Such a shift builds on previous work regarding normative "role beliefs" [@fitch_armed_1998] of the military institution, and has significant ramifications on the organizational culture of the military, its relationship with government and society, and American national security.

# The Field of Civil-Military Relations

Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and a number of other consequential issues. At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or how to ensure that the military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999]. Feaver describes this "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p. 214]. Within the context of the US, the study of civil-military relations often focuses on the relationship between military and government elites, or what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p. 380]. Indeed, it is at this level that vital decisions are made, including to go to war, how to prosecute war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be. A number of scholars have offered solutions to Feaver's "problematique," but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, often serves as the point of departure for many scholarly works on civil-military relations. Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's concept of normative civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes the concept of what Huntington called "military professionalism" [@huntington_soldier_1957].

In _The Soldier and the State_, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States. The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3]. The societal imperative, Huntington claimed, was a further function of two distinct features of the United States. The first was an "ideological constant," which referred to America's political ideology of liberalism [@huntington_soldier_1957, 143-162]. Huntington argued that liberalism as a political ideology stressed "individualism" [@huntington_soldier_1957, 90], and that American society was, since its founding, suspicious of standing military organizations and professional military forces. The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143]. Huntington claimed that the tension generated by these two imperatives - one functional stemming from external threats, the other societal stemming from the character of American political ideology and structure - needed to be solved. Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society. Both goals - military security and civilian control - had to be obtained.

Figure 1, below, shows a visual diagram of Huntington's logic in a political science framework [@feaver_civil-military_1996]. As Peter Feaver points out in a 1996 article, Huntington did not use such language when developing and explaining his concept. According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military. Huntington viewed the key independent variable that enables the achievement of these goals while also resolving the tension generated by the opposing functional and societal imperatives as the adoption of a type of civil-military relations framework which he calls, "objective civilian control" of the military [@huntington_soldier_1957, 83]. In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and behavior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.

```{r figure one, include=FALSE, echo=FALSE}
huntington_dag <- dagify(
    MSCS ~ CC,
    CC ~ FI + SI,
    SI~I + S,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MSCS", 1, .5,
      "CC", .7, .5,
      "FI", .5, .6,
      "SI", .5, .4,
      "I", .3, .3,
      "S",.7, .3,
    ))

huntington_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

huntington_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MSCS"="Military Security and\n Civilian Supremacy", 
                     "CC"="Patterns of\n Civilian Control\n 'Objective Control'", 
                     "FI"="Functional\n Imperative", 
                     "SI"="Societal\n Imperative",
                     "I"="Liberal Ideology",
                     "S"="Constitutional Structure")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory Of Civilian Control of the Military", subtitle="As Explained in The Soldier and The State, 1957")

```

Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83]. In fact, Huntington admits that the US Constitution actually envisioned and established a form of subjective vice objective scheme of civilian control of the military [@huntington_soldier_1957, 163-168]. What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats. It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general, that the military profession had emerged [@huntington_soldier_1957]. Huntington claims that Cold-War military officers, unlike the corporate body of officers that existed at the nation's founding, constitute a profession and thus, are part of a unique and distinct organization that differs from other occupations: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].

Numerous critiques surrounding the Huntingtonian conception of _military professionalism_ have been levied over the years. Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed. Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration - not segregation - of political and military activities. Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals. Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20].

Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020]. Risa Brooks charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs, which she alleges follows from a Huntingtonian conception of military professionalism, may develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17]. Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020]. Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that a significant issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020]. It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016]. In other words, the ground is always moving for the actors involved in civil-military relations.

## The Dependent Variable: Military Intervention in Politics and Military "Role Beliefs"

Given the normative model espoused by Huntington and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding why the military would intervene in politics. Numerous scholars, mainly writing from a comparative perspective (although not exclusively), have enriched our understanding of what possibly motivates the military to interject itself into the political sphere. To be clear, these scholars are focusing mainly on active-duty military organizations, and not on the specific block of retired military officers, as I plan to examine, but the motivations and mechanisms for intervention that they put forward enrich our understanding of the phenomenon of political intervention by the military.

The British political scientist, S.E. Finer, argues that the military's intervention in politics is a function of the opportunity and desire to do so [@finer_man_1962]. Finer identifies several factors which should normatively inhibit the military from intervening in politics. These factors are military professionalism, the principle of civilian supremacy, and fear that the military would fail to remain a cohesive fighting force should it fracture internally upon intervening in political matters [@finer_man_1962]. Finer critiques the concept of Huntingtonian military professionalism somewhat harshly, arguing that although military professionalism should act to focus the military on itself and thus not on political activities, those who claim to be professional might and have used the concept of professionalism as an excuse to not obey civilian authorities [@finer_man_1962]. Finer is clear that the principle of civilian control - not military professionalism - is the true and only bulwark that can keep the military out of politics.

Describing the various motives that militaries might have for intervening in politics, Finer goes on to describe at least three. He first lists the motive of the "manifest destiny of the soldiers," in which the military acts as "saviors" of the country [@finer_man_1962, 28-29]. This motive, Finer claims, arises because of the confluence of the military's unique mission, sense of power, and claim of being "outside party politics" [@finer_man_1962, 29]. A second motive that militaries might possess which disposes them to intervene in politics is what Finer calls the "unique identification with the national interest" [@finer_man_1962, 30]. Acting from such a motive, Finer conceptualizes the military acting as "custodians" of the "national interest" [@finer_man_1962, 30]. Finer acknowledges that "custodianship" can take on different forms in different contexts, but points out that some militaries might view "custodianship" as a "duty to arbitrate or veto...to exercise it when some convulsion or decision of the civil autorities seems to them [the military] to threaten what they think are the permanent interests of the nation" [@finer_man_1962, 31]. The final motive Finer spends the most time on, calling it "the sectional interest," which refers to a more narrow subset of interests than the "national interest," and includes a wide ranging set of interests based on social class, regional affiliation, personal career motivations, or organizational interests [@finer_man_1962, 34-53].

Ultimately, Finer develops a typology of four levels, or degress of, military intervention. The first he calls "influence," and argues that at this level of intervention, the military makes appeals "to reason or emotion" in an attempt to persuade the civilian government [@finer_man_1962, 77]. Finer adds that intervening at this level is, in his mind at least, not problematic and entirely "consistent with the supremacy of the civil power" [@finer_man_1962, 77]. The next level of intervention is "pressures or blackmail," describing efforts by the military to "convince the civil power by threat of some sanction" [@finer_man_1962, 77]. Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may be unproblematic, many other forms, such as the "threat to stage a coup," would be normatively wrong [@finer_man_1962, 77]. Finer points out a third level of intervention, that of "displacement," referring to the "removal" of government officials by the military as a result of "violence or the threat of violence" [@finer_man_1962, 78]. The fourth level, which is basically displacement on a wider scale, Finer calls "supplantment," which "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78]. Finer goes on to explain the level of intervention as a function of the level of "political culture" that is prevalent in a state and society  [do I need more here? See Finer, pp. 79 and following]

Another helpful conceptualization of military intervention in politics is the idea of "role beliefs," a concept articulated by the political scientist, J. Samuel Fitch, in his book, _The Armed Forces and Democracy in Latin America._

A third, and equally rich, study of the degrees of military intervention in politics is available from Brian Taylor's book, _Politics and the Russian Army: Civil-Military Relations, 1689-2000.

## Political Polarization

The question of if and how political polarization drives changes in the professional identity of the military spans several related literatures, most notable of which are American Politics and International Relations. Broadly speaking, this project is focused on how one aspect of domestic politics, political polarization, impacts a specific but important aspect of international relations, civil-military relations, which itself spans a wide and diverse literature. Because I intend to focus my work on the military as an institution, my dissertation will also rely on literature from the fields of sociology, organizational psychology, and ethics. I will frame my research question within each of these fields briefly, and to define important concepts as required, before addressing why this question matters and should be studied in depth.

Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], but there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012]. Despite the wide-ranging scholarship exploring the many facets of political polarization, virtually all scholars agree that contemporary levels of political polarization in the United States are historically abnormal and pose myriad problems.

One of the major concerns regarding political polarization is its on impact on American foreign policy in a broad sense, and specifically, on America's ability to lead and impact the world. Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017]. Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions will yield findings that may help policymakers and academics help solve the problems Schultz identifies.

## Military Professions and Organizational Loyalty

As a subject or concept, loyalty features prominently in the international relations and civil-military relations literatures, often in a comparative setting and in the context of coups [@marcum_overthrowing_2016; @johnson_squeaky_2018], civil wars [@mclauchlin_loyalty_2018], and the difficulties that multi-ethnic states face in maintaining a professional military [@peled_question_1998]. This literature deals with a host of important issues that stem from problems associated with ethnic identification and loyalty, either to a regime, cause, or leader. Though the US context is very different from the contexts usually explored in this literature, its insights may help me identify related dynamics and/or mechanisms that are unfolding in the US: how are changes in partisan or ideological identification (instead of ethnic) as a result of political polarization impacting conceptions of loyalty within the US military, if at all? To even begin answering this question, one has to first define what loyalty is and how it works. For this fundamental task, the fields of psychology, ethics, management, and sociology are helpful.

Loyalty has been conceptualized as both a virtue [@coleman_problems_2009] and as a moral emotion [@connor_military_2019], the main difference between the two being that the former is a normative aspirational trait, while the latter is a primed behavioral response. For the purposes of my dissertation, these differences are not consequential, because both assert that loyalty is something that is ultimately felt by an individual and subsequently reflected in action, such as, in the military context, "the desire to protect each other from harm and to defend each other from attack" [@coleman_problems_2009, 109]. Oldenquist, a philosopher, further describes loyalty as a devotion or attachment that an individual has to someone or something specific such that one refers to the object of loyalty as "mine" and is inclined to feel "pride when it prospers, shame when it declines, and anger or indignation when it is harmed" [@oldenquist_loyalties_1982, 175]. Loyalty also generates a willingness in the individual to assume "obligations" [@oldenquist_loyalties_1982, p. 175] on behalf of the object of one's loyalty, and these obligations can take multiple forms, such as a willingness of the individual to endure hardships and make sacrifices [@fielder_organizational_1992]. Scholars agree that individuals can hold multiple loyalties simultaneously, and recognize that this often can and does cause problems [@coleman_problems_2009]. For organizations, and by extension, professions, the concept of loyalty is also very important. Conner et al. state, for example, that "having loyal members is vital to perpetuating the group" [@connor_military_2019].

We also know that the military services in the United States place a premium on loyalty as a trait and concept. Indeed, of the different military services, the US Army in particular has extensively developed, in a normative sense, a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019]. This ethic, which will be described in detail in later parts of this dissertation, is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, and holds that as "stewards of the Army profession" and as "committed professionals," Army leaders "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019]. There are a total of seven "Army Values," which together spell the acronym, l-d-r-s-h-i-p, which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p. 1-9.].

ADP 6-22 further states that the Army Values "embody the practical application of the Army Ethic. They encompass the enduring moral principles, beliefs, and laws that guide Army professionals in accomplishing the mission as well as their conduct in all aspects of life" [@us_department_of_the_army_adp_2019, p. 1-12]. The manual defines loyalty as, "loyalty: bear true faith and allegiance to the Constitution of the United States, the Army, your unit and other Soldiers" [@us_department_of_the_army_adp_2019]. The virtue of loyalty is likewise extolled in a host of codes and mantras that have played a central part of the American military since its inception. In the middle of the West Point Cadet Prayer is the eloquently written plea, "make us to choose the harder right instead of the easier wrong" [@wheat_west_1918], indicating that to be loyal to what is right is not only preferable but often a difficult endeavor. Even Samuel Huntington recognized that in order for his model of civilian control to be effective, the military profession had to organize itself as a "hierarchy of obedience," which required "loyalty and obedience" to stand out as "the highest military virtues" [@huntington_soldier_1957, 73].

Loyalty is further important in a military setting in a broad and strategic sense, especially when one considers that the US military, both by statute and by normative standards, serves as institution with multiple, simultaneous loyalties. The Constitution of the United States, to which military officers and soldiers alike ascribe allegiance when they take their respective oaths of commissioning and enlistment, contrives the military as an institution possessing multiple loyalties: to the President as "the Commander in Chief of the Army and the Navy of the United States" [@noauthor_us_1787], but also to Congress in that they likewise hold unique powers and responsibilities pertaining to the Armed Forces [@noauthor_us_1787-1]. The tensions that inherently exist in American's constitutional framework with respect to national security are explored in a robust scholarly literature [for instance, see @howell_while_2007; @milner_sailing_2015].

# Where My Puzzle Fits and Why It Matters

In the context of the US military, to my knowledge, loyalty has rarely been used as a conceptual construct in the political science literature, likely and thankfully because the US military has not yet considered, or been involved, in a coup. Nonetheless, I contend that there is a rich opportunity to study strategic loyalty as a concept, especially in the contemporary context of the United States' domestic political climate, as political polarization has increased extensively.

One potential area of work that remains under-developed, both theoretically and empirically, is an exploration of the relationship between polarizing domestic politics and the professional identity and character of the military. I am specifically interested in the question of if and how political polarization impacts the concept of loyalty in the military. This is an interesting research question because it brings together at least three relevant strands of literatures which, when combined, generates an interesting set of puzzles and questions. First, one of the most fundamental normative hallmarks of the military's professional nature, at least by Huntingtonian standards, is that it ought to be an apolitical and non-partisan institution. Second, one of the features of the US military is that it prizes and elevates the trait of loyalty among those who serve and, by statute as well as normative tradition, exercises simultaneous dual loyalties to its civilian heads, the President as the Commander in Chief, and Congress. And third, we know from the American politics literature that ideological polarization and partisan loyalty are connected in that those citizens who hold to stronger ideological convictions tend to be the most loyal partisans in terms of voting and other political behaviors [@abramowitz_great_2018, 112-116].

If these things are true, then a question is whether rising political polarization will generate instances in which military officers must choose between which objects of loyalty to ultimately elevate and side with in the event that there is conflict between and among these objects. Moreover, at the level of the military institution, a similar problem may also emerge: will the military institution have to choose between competing objects of loyalty, such as the President as the Commander in Chief, the Congress, and even its own professional ethic?^[Loyalty had to be a prime consideration in several prominent, recent cases involving the military and the government. Consider the case of now retired Army Lieutenant Colonel Alex Vindman, who testified before Congress against the US President, Donald Trump, regarding what Vindman believed to be behavior by the President that may have been illegal, or the case of Navy Captain Brett Crozier, who leaked a letter to the press after his pleas for help in battling the coronavirus aboard the ship he commanded, the USS Theodore Roosevelt, went unheeded for what he felt was too long. Certainly these instances involved individuals making choices about to whom or to what they would be most loyal to - the President, the Constitution, the ideals and obligations of the military, the operational mission, themselves as individual moral agents, and perhaps even more objects of loyalty.]

Other questions I wish to explore include whether there patterns in the way that loyalty obligations held by the military are in fact challenged by governmental actors during periods of increased political polarization, and if so, what are they, and how is the military reacting to these patterns? Are there impacts on the military's professional identity as a result of political polarization, and if so, from who or what? Are partisan/ideological loyalties displacing other loyalty relationships within the military?  Through what mechanisms do these changes operate? What sorts of changes are they driving, if any, in the military institution? Is this problematic for the military profession, and why? These and other questions will shed light on how rising political polarization is impacting the professional identity of the military.

The gaps in the current literature, which I hope to advance and fill, involve the impact of political polarization directly on the military institution itself, and why the military experiences such impacts. A number of excellent and recent scholarly works have looked at aspects of this question, but from an angle that examines not the military directly, but rather society's perception of the military or various aspects of the military as a result of polarized politics. A number of these works have directly examined the impact of partisan affiliation on the credibility of military actors, resulting in some very interesting findings. Michael Robinson, for instance, found that members of the public did not sanction military elites for making public messages that accorded with a member's partisan views, and that the public only sanctioned those elites whose messages were in line with the opposite political party [@robinson_danger_2018]. Previously, Heidi Urben found that among active military officers, there was weakening support for the norm that retired military officers should remain apolitical or non-partisan in retirement [@urben_party_2013]. These scholars' findings suggest, as others have pointed out [@brooks_paradoxes_2020], that political polarization at least coincides with, if not causes, a weakening in support among both the public and among service members for the military's normative apolitical ethos. Other scholars have also unpacked a trove of information regarding public opinion and the military, including how partisan affiliation shapes differences in support for the use of military force [@golby_jim_elite_2017] and military elite endorsement of political candidates [@golby_military_2012].

Focusing attention on the military institution will help scholars in the field of civil-military relations and international relations more broadly better understand if and how the military, as a unique body charged with the defense of the nation, changes in fundamental ways as a result of a polarized political climate. On one hand, one would expect that it might, in the sense that the American military is, after all, a microcosm of American society. Therefore, one could expect that if society truly is becoming polarized, then at least to some degree, so is the military, or at least those who volunteer for service and at the point in time that they join. On the other hand, one might reasonably expect that given the unique purpose of the military, combined with its long-standing professional ethic, might insulate the military institution from some of the potentially negative impacts of polarized politics. These differing expectations need to be more clearly articulated and explored.

To be clear, scholars have long studied how domestic politics impact the military. And as scholars have pointed out, military and civilian relationships are in tension by their very nature in that the leaders of a democracy allow a most undemocratic institution, the military, to protect the polity [@feaver_civil-military_1996]. And as American history proves, there have been countless instances in which civilian and military leaders disagreed. Many of these rise to the level of common knowledge: Lincoln fired McClellan because he lost confidence in him as a wartime Commander; Truman sacked MacArthur because the latter was insubordinate regarding the future of war in Asia; and Obama and McChrystal clashed about troop levels in Afghanistan, ultimately resulting in the former firing the latter. Beyond these episodes in which President and wartime commander locked horns, consider several other instances of "civilians" and "military" clashing. One episode is certainly the way in which the Vietnam war escalated in ways that, at least according to H.R. McMaster, military generals at the time knew were destined to fail [@mcmaster_dereliction_1998]. A more recent example, and one much studied and commented upon by civil-military relations scholars, is the so-called "Revolt of the Generals" that occurred following the early years of the invasion of Iraq, which involved several recently-retired generals and admirals sharply criticizing Secretary of Defense Rumsfeld's leadership dynamics [@cook_revolt_2008].

My intended research topic is different than these other, still very important topics in which disagreements between the executive and the military occur during wartime. It is my assessment that polarization is causing conflict, or certainly has the potential to do so, between the military and civilian leaders, but in most instances, the subject is about topics other than the use of military force, which is somewhat puzzling, given the military's reason d'etre. For example, I believe we are seeing portions of the military institution (particular retired military officers) openly disagree with civilian leaders on topics that I believe are, even by the widest of academic definitions, outside the bounds of what constitutes appropriate military expertise.^[I hope to prove this, empirically, in or as part of a chapter in my dissertation]

This raises, then, an important question: does the military's essential purpose risk evolving or changing significantly during polarized times? While I believe that these potential developments are catching the attention of civil-military relations scholars, none to my knowledge have specifically set out to systematically explore if and how polarized partisan politics are impacting the military's professional identity and character, and in particular, the role of the military institution itself in facilitating these developments. Most scholars acknowledge the potentially damaging impacts of the politicization of the military [@golby_listening_2013; @brooks_paradoxes_2020] at the hands of civilian leaders, and I share these concerns. Fewer scholars, if any, are exploring the potential role of the military in politicizing itself, and the mechanisms through which these developments occur.

# The Analytical Framework - Initial Argument

## Political Polarization Impacts Conceptions of Loyalty

In this section of the prospectus, I will lay out an initial argument, as well as accompanying hypotheses, which I intend to explore and test in my dissertation. Figure 2, below, shows an initial conception of the linkages between several of the variables discussed in the previous section of the prospectus. From left to right, this conceptual model suggests that worsening political polarization is driving changes in both the institutional credibility of the military (how society views the military) and changes in the way members of the military institution think about and exercise loyalty, and by extension, how they think of their professional identity. These two independent variables are, I suggest, driving subsequent changes in the concept of military professionalism, the main dependent variable in my analysis.

In line with Huntington's model as well as a host of other scholarship, I suggest that changes in military professionalism can impact a variety of military outcomes, varying from effectiveness on the battlefield [@reiter_democracies_2002; @biddle_democracy_2004] to achieving strategic effectiveness [@brooks_creating_2007; @brooks_shaping_2008]. While I intend to examine the broad system of variables shown in Figure 2, the focus of my dissertation is the relationship between political polarization and changes in the concept of loyalty and identity, and how such changes manifest themselves in the variation of military professionalism. For emphasis and clarity, I do not think that the other link - that between political polarization and institutional credibility - is unimportant. Rather, this link seems to have been the focus of several recent scholarship efforts [@robinson_michael_danger_2018; ].

```{r figure two, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    MO ~ MP,
    MP ~ LI + IC,
    IC~PP,
    LI~PP,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MO", 1, .5,
      "MP", .7, .5,
      "IC", .3, .3,
      "LI", .3, .7,
      "PP", .1,.5,
    ))

problem_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MO"="Changes in\n Military\n Outcomes", 
                     "MP"="Military\n Professionalism", 
                     "PP"="Increasing\n Political\n Polarization", 
                     "IC"="Institutional\n Credibility",
                     "LI"="Concepts of\n Loyalty and Identity")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="The Impact of Political Polarization on Military Professionalism")

```

A primary contention from the system visualized above is that as political polarization in the domestic sphere increases, the military faces increased challenges in how both individual military leaders as well as the institution conceptualizes, adopts, and exercises loyalty, and by extension, how it sees its own professional identity. This expectation stems from several features that scholars of American politics have raised regarding the current politically polarized climate within the United States. First, scholars have noted that political polarization raises the number and scope of issues that are subsumed by partisan debate and disagreement (a phenomenon referred to among scholars of American Politics as "conflict extension" [@layman_party_2002]). A second feature of contemporary political polarization is the prominence of  negative feelings held by one partisan group towards another, a phenomenon referred to as "negative partisanship" [@abramowitz_great_2018]. A third feature is an increase in the salience of debates over issues that individual partisans consider core to their individual identities [@white_values_2002; @mann_its_2012; @sides_identity_2018]. I contend that these features impact individuals serving in the military, including military leaders, as they do members of the domestic public writ large, and that these impacts are problematic for members of an institution that normatively exists as an apolitical and non-partisan actor. Table 1, below, lays out how these three prominent features of domestic political polarization may impact the non-partisan and apolitical nature of the normative military professional and institution.

\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{How Features of Contemporary Political Polarization Impact Normative Military Professionalism} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Professionalism}  \\ \hline
    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number/scope of issues. & Military Professionals and the Institution find it increasingly difficult to "stay out" of partisan politics as the territory of what is considered partisan politics grows. \\ \hline
    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain arise against those who hold to opposing political views. There is a belief that those on the other side of the political aisle are not just mistaken over policy, but greatly wrong and flawed as individuals. & Even among military leaders, there is a greater willingness to confront or take issue with those who hold opposing views.\\ \hline
    \textbf{Individuals believe that "core" issues are increasingly at stake} & The salience of one's identity and core values is raised as partisans believe essential issues are at stake in the political arena. Associated with this is a belief that not just policy preference, but moral right and wrong, are at stake. & Military professionals, who hold multiple institutional loyalties, face an increasingly difficult task of balancing these loyalties, especially as the objects of these loyalties are increasingly associated with or attached to notions of moral right and wrong. \\ \hline
\end{tabular}
\label{table:comparison} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

The first way in which the apolitical and non-partisan ethos of the military is challenged as a result of increasing political polarization is that there is increasing difficulty for individuals to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of military professionalism. This happens because as "conflict extension" [@layman_party_2002] occurs, an increasing number of issues are brought into partisan battlegrounds, and thus, there is simply less politically-neutral territory available. Thus, an actor such as the military, who may wish to remain non-partisan and out of politics to the extent that it can, may simply be less able to do so.

A second challenge military leaders and the institution faces in a polarized climate is from the rise of "negative partisanship"[@abramowitz_great_2018] that dominates the polarized American political landscape. As citizens hold increasingly strong and negative feelings towards those who side differently on a host of issues, there may be a tendency for members of the military institution to harbor similar feelings and attitudes. Even if the impact of this feature on the military and its leaders, due to training and professional development that emphasizes the importance of upholding apolitical and non-partisan behavioral norms is somewhat muted, the point is that members of the military institution are likely experiencing the same phenomenon - even if it is to a lesser degree - as the country's citizens. One important question about "negative partisanship" in a military setting, though, is whether service members hold strong negative feelings toward only civilians who are on the opposite side of the ideological spectrum, or whether such feelings also permeate within the ranks.

A third, and I contend the most impactful, challenge that normative military professionalism faces stems from the belief by individuals that core issues - such as those that involve attitudes about race, religion, and multiple layers of identity - are at stake in the current polarized political climate. "Core" issues are difficult to define precisely, but the main point is that these issues often involve perceptions of identity [] and moral right and wrong, and not just simple policy preferences. Moreover, some of the most polarizing issues are often associated with issues related to an individual's identity, such as race, gender, immigration, as the the rise of "identity politics" in America illustrates this point [@sides_identity_2018].

I argue that this third challenge is especially relevant for the contemporary military, and specifically, with respect to how members of the military exercise the concept of loyalty. As an increasing number of what individuals believe to be core issues are raised and debated in the partisan arena, individuals face decisions of how to prioritize the persons or objects of their loyalty and further, what actions to take on the basis of loyalty to these persons or objects. I further theorize (though this has to be tested, of course) that while all individuals will likely face some version of this challenge of loyalty in a polarizing age, the challenges will be more acute for military leaders. Not only is the de-facto setting in which military leaders operate marked by multiple and often conflicting loyalties due to the structure of and statutes set forth in the US Constitution, but also, the potential consequences are higher for military leaders than average citizens should they fail to discharge the duties that correspond to their roles as professional officers. These consequences include failure to achieve tangible national security goals, such as loss on the battlefield, or less tangible but still important outcomes such as maintaining trust with the American public.

As important as these consequences are, they do not speak to another set of consequences individual military leaders also may face in a polarizing age. Indeed, in addition to consequences to the country that may occur should the military fail to discharge its duties to safeguard America, military leaders also potential consequences if confronted with a situation in which they have to execute orders which are technically legal, but morally borderline at best. Military leaders are still "moral agents" who "remain responsible for the decisions and actions they take" [@dubik_just_2016, 160]. But the considerations of what constitutes moral right and wrong to begin with may be changing as polarization worsens [@graham_liberals_2009].

Therefore, I argue that if polarization continues to heighten the identities of groups that people associate with, as well as come closer to the boundaries of how individuals think about moral right and wrong, there is an increased likelihood of military leaders having to choose between conflicting loyalties. What happens if and when the objects of a military leader's multiple loyalties not merely disagree, but become stubbornly divided ideologically? Can such an outcome exist? Is there empirical proof that such a trend is happening? If so, what are the normative implications, both for the military as a profession, and for the nation's security, and are there actions that can be taken to halt such a trend?

## How Polarization Interacts with Institutional Credibility

The previous section described how increasing political polarization presents significant challenges to aspects of the normative concept of military professionalism. But whether these challenges enter the public domain, or remain private challenges that specific individual actors contend with on a personal level only, matters a great deal because it "changes the character...from a matter of principle to a political matter" [@dubik_principled_2014] In fact, in much of the field of civil-military relations, this dynamic of private disagreement or discontent potentially entering the public realm is seen in the scholarly literature on senior officer resignation. This debate examines various aspects of resignation, including whether there can ever be conditions under which resignation is justified or appropriate, with scholars taking various sides on the issue, both for [@dubik_principled_2014; @snider_strategic_2014] and against [@feaver_should_2014; @golby_beyond_2015]. I contend that polarization and credibility of the military interact, however, in at least two important ways. First, as loyalties are challenged as a result of increasing political polarization, individual military leaders may not necessarily remain quiet despite the existence of a strong professional norm of doing so because the objects of loyalty may demand that their adherents publicly speak out. This of course, depends specifically on what the object of loyalty is and what it demands from those who hold closely to it. The second way in which polarization and credibility may interact arises more directly from what can be termed a credibility effect, and has more to do with the place of prominence (or lack thereof) that the military or individual leaders hold in the perception of American society. As recent scholarship has shown, the credibility of the military institution in a variety of contexts depends heavily on partisan affiliation [@urben_party_2013; @golby_listening_2013; @golby_jim_elite_2017; @robinson_danger_2018].

I contend that when institutional credibility is high, a demand signal is created by the American public such that future interactions between the military institution and the American public are encouraged and welcomed. From an institutional perspective, the military might respond to the public's demand signal for increased military interaction by increasingly participating in public ceremonies, parades, and other showcases of military prowess (at sporting venues, for instance). From an individual perspective, military elites, especially retired military officers, would be psychologically primed to increasingly dialogue with the American public when military credibility is high. By extension, other outlets such as the press would also be primed to showcase military elites when military credibility runs high, because these elites represent an institution that commands a high degree of public respect.

When military credibility is low, the theory contends the opposite will occur. Since demand for interactions between the American public and the military is muted when institutional credibility for the military is low, the institution will not engage with the public as frequently nor as intimately because the strong demand signal that existed when institutional credibility was high no longer exists. At the individual level, military elites will also not engage as often with the public, nor will the press ask these elites to. In sum, institutional credibility and respect for the military institution impacts the likelihood of the military institution and military elites interacting with the American public. Such interaction can take a variety of forms. When respect for the military is high, the military is likely to interact, speak to, and dialogue with the American public, and when institutional credibility is low, the military is less likely to do so.

Ultimately, four possible worlds exist when the impacts of military credibility and political polarization are considered in tandem (see Table 2, below). The first is when political polarization and military credibility are both low. When these conditions hold, there is no anticipated drop in military professionalism by the military institution nor its individual elites. With these conditions, the military does not have an out-sized platform to interact with the American public, nor are there outstanding incentives for the military to raise significant disagreement over any domestic political issues.

\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{Four Types of "Military Professionalism Worlds" Where Institutional Credibility and Political Polarization Vary} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    Influence of Variables & \textbf{Low Institutional Credibility} & \textbf{High Institutional Credibility}  \\ \hline
    \textbf{Low Political Polarization} & Negligible anticipated impacts. Traditional Professionalism is Expected by the Military, Civilian Leaders, and the American Public & Possible Impacts. There is a demand for military voice, but little substantive matter stemming from US domestic politics that can cause crucial disagreement. Little incentive to deviate from traditional concepts of loyalty and duty. \\ \hline
    \textbf{High Political Polarization} & Possible Impacts. Polarization raises the stakes of domestic politics, but military professionals are constrained from voicing their opinions and/or interfacing with the public due to low institutional credibility with the American public &  Highly Likely Impacts to Professionalism. Likely impacts. Military professionals have a "platform with the public" due to high institutional credibility, and high polarization raises the likelihood of disagreement between individual military elites and civilian leaders on a number of crucial topics.\\ \hline
\end{tabular}
\label{table:comparison1} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

The second possible set of conditions is when political polarization is low and institutional credibility is high. When these set of conditions exist, the military has a platform to interact with the public, as the public highly respects the military institution. The military enjoys this high level of trust and does not want to lose it. Further, because the degree of political polarization is low, there is little incentive for the military as an institution or military elites to use the trust they enjoy with the American public to adopt or denounce an issue associated with any partisan entity. The high institutional credibility increases the chance of the military interacting with the public in an unprofessional way, but the low political polarization removes the incentives of doing so.

The third type occurs when political polarization is high, but institutional credibility of the military is low. Under these conditions, the military does not command the respect of the American public like it does when credibility of the institution is high, and thus, does not possess a considerable outlet to interact with the American public. However, the incentives of loyalty and duty still loom large in the minds of some military elites as a result of the high degree of political polarization, and thus, a dip in professionalism is still possible.

The final set of conditions exists when political polarization and institutional credibility are high. It is in this world that I contend the American military finds itself in now in 2020. Under these conditions, the military faces dual dangers. High institutional credibility risks the military institution unilaterally acting or another body co-opting the military institution to speak out an a number of issue areas, and simultaneously, the highly polarized environment risks the military advocating for or denouncing partisan sentiments in the name of upholding the very virtue it is most normatively adhered to - loyalty.

# The Analytical Framework - Hypotheses

In order to test my argument, I will have to examine and test several hypotheses, which are inter-related and linked. They include the following:

\singlespace

_H1: In highly polarizing times, there will be an increase in the frequency of military officers breaking normative civil-military relationship behaviors, such as criticizing the President, members of Congress, or those who hold to different political views on particularly-polarizing issues outside of the conduct of war._
\doublespace
\singlespace

_H2: These breaks in normative civil-military behaviors will occur because political polarized periods feature events or situations which test or challenge the military's ability to juggle and balance multiple simultaneous loyalties: to the President as Commander in Chief, to Congress in its oversight role of the military, to the underlying ideologies or ideas inherent in issues that prevail during polarized times, to the normative professional ideals of the military, and to the individual's bonds of loyalty to his or fellow soldiers and the mission._
\doublespace
\singlespace

_H3: Because of the normative strength and prominence of place that the ideal and concept of loyalty has within the profession of arms, military leaders are more sensitive than regular civilians to appeals of loyalty._
\doublespace
\singlespace

_H4: Following from the above hypothesis, polarized periods will feature events in which governmental actors appeal to the military's sense of loyalty in order to persuade them to adopt a stance on a particular issue area. These can be viewed as "exploits of loyalty."_
\doublespace
\singlespace

_H5: In highly polarized time periods, it is the senior military officers as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p.1-9] whose actions will most influence to what degree and how the military institution reacts to and changes as a result of "exploits of loyalty." If changes occur to the military institution, they are bound to occur to the professional identity of the military, which includes the degree to which it functions as a profession (vice bureaucracy [@snider_once_2012]) and the degree to which it is obedient to its civilian leaders (vice acting as an activist on behalf of a particular cause)._
\doublespace

# Plan of the Dissertation and the Data Generating Process

To complete my dissertation, I plan on incorporating a mixed-methods approach. In broad conceptual terms, I plan to propose a theory and then test the theory using a combination of quantitative analyses and qualitative case studies. The one sentence thesis of my dissertation is that political polarization significantly impacts how the military thinks about and exercises loyalty, a chief military trait, and that these impacts can and often do result in changes to the military's professional identity and character. I plan to execute a book style project comprised of seven chapters.

In the first chapter, I will frame my research question, lay out my research agenda, and review the relevant literature for my project. This will involve synthesizing several of the key aspects and debates within the field of civil-military relations, much in the same way I have done in the first part of this dissertation prospectus. I will focus my attention on introducing and summarizing the concepts of military professionalism, organizational loyalty, and political polarization, and spend time orienting the reader to the debate within scholarly literature regarding a Huntingtonian conception of military professionalism. I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020]. The main point of chapter one is to set the stage for my dissertation and signal to the reader that modern critiques of Huntington may be overlooking a key part of his concept of military professionalism: imperfect as it may be, it may also insulate the military from the most pernicious effects of polarizing politics.

In the second chapter, I will introduce, develop, and explain my theory that links increasing political polarization to conceptions of loyalty and military professionalism (see Figure 2, above). I will argue that specific features of the polarized environment strain and challenge both individual and institutional conceptions of loyalty, and by extension, normative aspects of military professionalism (Table 1). I will further argue that rising political polarization can interact with changes in the military's institutional credibility to increase the likelihood that the public becomes aware of how the military chooses to respond to the dynamics brought on as the military faces challenges to how it conceptualizes and practices loyalty. Using Table 2 as a framework, I will place epochs of historical American civil-military relations into one of four distinct quadrants, and signal that these quadrants will comprise the qualitative case studies that I later examine in my dissertation.

In the third chapter, I will argue that today, in 2020, the American public is witnessing a decline in professional military conduct among retired military officers. I plan to demonstrate this empirically by drawing upon an original data set consisting of more than 340 op-ed pieces authored by retired military officers since the late 1970s and published in the top five major print newspapers in the United States, including the Wall Street Journal, the New York Times, the Washington Post, the USA Today, and the Los Angeles Times. I believe that I can show several things from this analysis, including that there has been a sharp uptick in the number of op-eds published by retired military officers during this time period, and more relevant to the theory I propose in the dissertation, that these published op-eds reveal a decline in the normative professional ideals of the military institution. Unlike previous recent works which focus on how polarization impacts society's view of the military, this chapter introduces an empirical record of actors within the military institution (retired senior officers) also changing as a result of increasing political polarization.

Critical to this chapter is acknowledgement that coding for what is professional versus what is not is a somewhat subjective process. Thus, I will need to develop and implement a clear working definition of the concept of military professionalism in Chapter 1. However, I believe that I can show in Chapter 3 that whether one adopts a strictly Huntingtonian version of military professionalism (strict separation of political and military spheres) or a less strict version that allows for an increasing shared space, the retired military officer corps is not living up to the non-partisan ethic that the military espouses.

I will argue that while there are certainly other confounding variables, to include changes in the conservative-liberal tilt of newspapers over time, the chief explanation for the variation (decline) in professionalism stems from the combined effects of increasing political polarization and rising institutional credibility of the military since the late 1970s, and the challenges that these changes introduce to a military leader's conception of loyalty. Moreover, my ultimate point is to point out a blatant inconsistency in the way that the military currently thinks about the political behavior by those within the ranks in that the institution has rules that governs the conduct of those on active duty [@department_of_defense_political_2008], but those same rules, and the spirit behind those rules, are disregarded by the institution's vanguards, its retired multistar officers and "stewards of the profession" [@us_department_of_the_army_adp_2019, p. 1-9] of arms. I believe that this chapter will result in finding support for _H1._

In Chapter four, I plan to analyze the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers (pending a lengthy approval process through the Army). I intend to design a survey experiment that measures whether arguments that appeal to one's individual or collective sense of loyalty are more effective than those messages that do not incorporate such language, and more directly, whether such arguments are acutely effective among military audiences. Such an experiment will build off of recent surveys that have examined the impacts of "moral reframing" [@feinberg_moral_2019] and "moral messaging" [@luttrell_challenging_2019], which describe efforts senders of messages employ to persuade their audiences by appealing to the audience's sense of morality.

Moral reframing is an extension of the broader _Moral Foundations Theory_, a paradigm developed by psychologists several years ago [@graham_liberals_2009] that seeks to explain the moral domain in terms of five distinct moral foundations: "harm/care, fairness/reciprocity, ingroup/loyalty, authority/respect, and purity/sanctity" [@graham_liberals_2009]. Several follow on studies have confirmed that political liberals and conservatives rely on these five moral foundations in profoundly different ways, particularly when making "moral judgements" [@graham_liberals_2009]: liberals tend to rely on a morality based primarily on the foundations of harm/care and fairness/reciprocity, while conservatives tend to rely on a "more even distribution" of the five values, to include the "binding" foundations of ingroup/loyalty, authority/respect, and purity/sanctity [@graham_liberals_2009]. My intended survey experiment will advance the field because it will explore the degree to which one particular moral foundation - ingroup/loyalty - resonates within an organization that emphasizes and values this particular moral foundation, as I have laid out in the preceeding pages. To my knowledge, while there have been many studies that examine the extent to which partisans rely on differing moral foundations [need cite], there have not been studies that examine the degree to which one particular moral foundation binds members of an organization. The military serves as a potentially very interesting case study, particularly concerning the moral foundation of ingroup/loyalty.

Therefore, I will test whether an appeal to loyalty works stronger among members of the military than among members of the civilian population, as I have hypothesized in _H3._ The survey, after controlling for gender, partisan identification, and age (experience for military), will present the respondent with 6-8 hypothetical situations and a corresponding argument made by military voice (the biographical particulars of which I will also vary). Members of the treatment group will receive an appeal to loyalty as part of the military voice's argument as to why a certain course of action should be followed regarding the hypothetical situation presented.

The survey will be structured such that the scenario and corresponding course of action contain references to issues that are, at least in some of the cases, highly correlated with partisan / ideological affiliation. I will also vary the treatment such that while each treatment will be an appeal to loyalty, the specific object of loyalty within the treatment condition will be varied. For example, some questions will appeal to the military member's sense of loyalty to the Constitution, while another may appeal to the military member's sense of loyalty to the mission and/or the member's fellow soldiers. An example of one of these questions is as follows:

\singlespace
* Situation: The President of the United States, a [Republican, Democrat] recently commented that he is seriously contemplating ordering active duty military Troops to assist local police and National Guard units in several US cities in maintaining law and order as a result of initially-peaceful protests which turned violent in some, but not all, cases. These protests have been going on for more than two weeks and are largely, but not exclusively, affiliated with [the Black Lives Matter Organization, a pro-2nd Amendment organization].

* Elite Voice: A retired US Army 4 star general, who has extensive combat experience in Panama, Iraq, and Afghanistan, recently penned the following op-ed piece in a prominent newspaper:

* “The military must not be ordered to put down riots and protests, even if a small number of them have, unfortunately, turned violent. Even if the President is legally able to issue such an order, the military’s fundamental purpose is to defend the nation from external threats, and not to endanger the Constitutional rights of everyday Americans. [Treatment group, add ]: Should the President order the active military to deploy to cities where these protests have turned violent, military officers should resist such an order and if necessary, resign their commissions. They owe a higher loyalty to the Constitution, and the President’s order would clearly violate the oath of office these officers have sworn to uphold.”

* Do you find this argument compelling (7 point scale of (1) not at all - (7) absolutely, very compelling).

\doublespace

Such a survey will enable me to measure at least three things. First, it will allow me to measure whether appeals to loyalty are truly more effective than standard arguments that do not invoke obligations of loyalty. This result can be measured by comparing the answers between the treatment and control groups for both populations. The second object this survey will enable me to measure is whether members of the military (or veterans) react differently to appeals based in the trait of loyalty than regular civilians, as I have hypothesized in _H3_. The third thing this survey will allow me measure is the relative strength, if any, of appeals made to differing objects of loyalty (to the Constitution, to the Commander in Chief, to the sense of mission, to one's fellow soldiers, etc.) in relation to a particular course of action. The results of this chapter will inform a micro-foundational understanding of the strength of loyalty as it pertains to the decision-making of individuals.^[Of particular note, I am planning to also pursue a survey among actively-serving military officers, but the lead time for obtaining approval to do this can be very long, perhaps 9-12 months or more. If I am able to obtain approval for this, I will then have three different populations to measure: civilians, veterans, and actively-serving military officers.]

In the fifth chapter, I want to show that the military has also faced institutional challenges to the concept of loyalty as political polarization has worsened. To do this, I plan to examine, via textual analysis, the records of Congressional Committee hearings in which military officers, active and retired, have participated. My goal is to detect if and how patterns of questioning and providing answers have changed, particularly in how language related to the concept of loyalty has been used. I still have some work to do in fleshing this out some more, but my initial idea is to examine how the military is both asked and answers questions by members of Congress, and how these questions and answers invoke language that speaks to various loyalty obligations. An example of this might be analyzing the ways in which terms such as "moral obligation" and "sacred duty" have been used over time. I might be able to associate causes or subjects that are associated with this type of language, and correlate these causes and subjects to the level of polarization present in the country, as well as to the partisan identification of the specific member of Congress asking the question. My goal is to detect patterns that will help support my overall argument, and specifically, _H2_.

In the sixth chapter, I will turn to a series of qualitative case studies to explore and test my theory. Key in this chapter is finding and examining historical periods in American civil-military relations in which there existed varying levels of political polarization and institutional credibility, and then, to walk the reader through how challenges of loyalty during these periods manifested themselves in different ways among specific members of the military and the institutional military at large, in line with the theory that I propose in Chapter 2 and specifically, _H2_.I then hope to find evidence for how the actions of the military and its leaders during these periods impacted the military's professional identity and character, in line with _H4_ and _H5_.

I have not yet settled on what the specific case studies be, particularly because I am wresting with whether I need to confine my examination to the time period in which the US military has generally been considered a profession (depending on who you ask, this is generally around World War One, or perhaps a little bit earlier, at the tail end of the 19th Century). My initial reaction is that I do not need to confine myself to this period. My initial projection of what these case studies will be and how they will align with varying levels of political polarization and institutional credibility are as follows:

* High Polarization, High Institutional Credibility: civil-military relations in the Andrew Johnson administration after the American Civil War. I will explore how General Ulysses Grant confronted challenges of loyalty to multiple institutions and objects as he dealt with a President who was by many accounts sympathetic to southern interests as the US Army played a prominent role in reconstruction in the south [@hyman_santon_1960].

* High Polarization, Low institutional credibility: civil-military relations after World War One, during the 1920s.

* Low Polarization, High Institutional Credibility: civil-military relations after World War Two, in the late 1940s and early 1950s.

* Low Polarization, Low Institutional Credibility: civil-military relations in the post-Vietnam years, roughly 1972-1980.

In the seventh and final chapter, I will conclude by exploring the implications of my argument.  Note: one idea I am considering is doing an extra chapter or at least an extended portion of Chapter 6 on how the Black Lives Matter movement is impacting the military profession.  I think there could be some interesting work here.

# Initial Concerns and Timeline

My main concern at this point is that I harness the relevant outside literature, from the fields of psychology, ethics, sociology, and management, and still focus my puzzle on an empirical, political science question. To that end, I am concerned that I have a topic that is not sufficiently narrow, and thus, I may need to hone in on a specific subset of questions that I have presented here.

A second concern is that I do my best to make this dissertation relevant and interesting for international relations, and not only for the military itself (though that is certainly a focus of my intended work). To that end, I may need to adopt a more comparative approach that examines if and how other militaries have been impacted by political polarization in their countries, and thus attempt some sort of cross-national study, should one of my intended chapters be deemed insufficient. I note this especially because the concept of loyalty, as mentioned in this prospectus, features so prominently in international relations in the context of coups, ethnic exclusion, etc., and I want to make sure I am not missing an opportunity to bring these matters to bear on my research project.

In terms of timeline, and this is still very fluid, I intend to reach the following milestones:

* Between now and 15 December 2021: Continue to work on my prospectus. Work with Jonathan Renshon and Jon Pevehouse to develop and field my intended survey experiment (I have already obtained IRB approval through UW-Madison, but I need to apply for funding (through the Army), finalize the experiment, and prepare for data collection.) Plan to take my methods (second field) comprehensive exam.

* No later than 15 January 2021: defend my dissertation prospectus.

* No later than 15 May 2021: fully write Chapters 1-3 of my dissertation and submit for review to the dissertation committee. Prepare to field the original survey experiment for Chapter 4. Begin data collection for Chapter 5 of the dissertation, and settle on which case studies to write for Chapter 6 of the dissertation.

* Summer 2021: continue working on Chapters 4 and 5 of the dissertation.

* Fall 2021: submit drafts of Chapters 4 and 5 for review to the committee, and begin writing Chapters 6 and 7.

* January 2022: submit initial drafts of Chapters 6 and 7 to the committee for review.

*February - May 2022: improve all chapters of the dissertation.

* No later than May 2022: defend dissertation.

# Conclusion

The purpose of my dissertation is to explore how polarizing domestic politics affects normative concepts of military professionalism, particularly in the context of the United States and among the leaders of the US military. Certain features of increasing political polarization - in particular, the way in which individuals associate their partisan beliefs with salient features of identity [@sides_identity_2018] and right and wrong [@graham_liberals_2009], I argue, introduce immense challenges to the way military leaders conceive of and exercise loyalty, which has long been a foundational trait of military professions [@us_department_of_the_army_adp_2019]. In particular, military officers in polarized time periods face challenges of conflicting loyalties [@coleman_problems_2009] that are more acute than during non-polarized periods, resulting in situations where military officers must prioritize the multiple objects of their loyalty (to the President as Commander in Chief, to Congress, to the operational mission, to subordinates and superiors along the chain-of-command, and to one's self).

These decisions, though made by individual military leaders, often result in changes to the military as a profession and as an institution, with ramifications that bear on civil-military relations, the military's ability to defend the nation, and other important areas. By focusing my dissertation on this topic, I will contribute to scholarship by examining the specific ways in which polarization impacts the military institution, thus helping scholars of civil-military relations and international relations better understand the mechanisms through which polarization changes American institutions, particularly among the one institution charged with maintaining national security.

\singlespace
\pagebreak