### 2020-12-04 code/prospectusv2.Rmd

---
title: "Servants, Syndicates, and Saviors: The Rise of Domestic Political Interventionism by Senior Military Leaders (Prospectus)"
author: "Peter M. Erickson"
date: |
  `r format(Sys.time(), '%B %d, %Y')`
abstract: |

  _This dissertation explores changes in how the contemporary military institution intervenes in domestic politics. I advance a theory which argues that normative beliefs of military officers regarding political activism is a result of two interacting factors: the level of domestic political polarization within a society, and the degree of prestige and credibility the military holds. To support my argument, I trace changes in levels and types of distinct forms of political activism by military officers, including authoring opinion-editorials in major newspapers, as domestic polarization and military institutional credibility rises. I ultimately argue that the increasing political activism of military leaders is not merely a response to the hyper-partisanship present in the current political atmosphere, but rather the logical outworking of trends involving the prestige of the military institution as well as political polarization that have been building for decades. I conclude by describing the implications of changing beliefs among military leaders regarding the role the military should have in domestic politics. This dissertation contributes to the fields of international relations and civil-military relations by advancing an empirical and historical analysis of a growing phenomenon: military leaders intervening in domestic American politics. _

geometry: margin = 1.15in              # margin widths
fontsize: 12pt                         # text size
bibliography: researchpaper-ps904.bib # bibliography file
biblio-style: authoryear               # citation style
urlcolor: blue                         # links to external URLs
citecolor: black                       # links to bibliography
linkcolor: magenta                     # internal figures, footnotes, etc.
indent: true                          # paragraph format
                                       # true for indented paragraphs

header-includes:
    - \usepackage{setspace}
    - \usepackage{multirow}
    - \usepackage{indentfirst}

output: 
  bookdown::pdf_document2:
    keep_tex: true
    latex_engine: pdflatex
    number_sections: false
    toc: false
    fig_caption: true
    citation_package: biblatex
    includes: 
      in_header: 
        - latex-preamble.tex
  bookdown::htmldocument2: default
  bookdown::word_document2:
      toc: false
---

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, echo = FALSE}
library("here")       # file paths
library("tidyverse")  # workhorse package
library("tidylog")
library("kableExtra")
library("knitr")
library("ggdag")
library("dagitty")
library("gridExtra")
library("broom")
library("dplyr")
library("reshape2")
```
\newpage

# Introduction and Research Question

Across the world and in almost nearly every country, discussions regarding the participation of the military in politics are nothing new. Indeed, few topics evoke such strong opinions and emotions than whether and to what degree the military, both as a corporate body and as individual military officers, should be political vice remain separate from certain aspects of and activities associated with politics. These normative questions rightly form a significant portion of the study of civil-military relations, and intellectuals, practitioners, and pundits have frequently weighed in on these topics over the years.

Thankfully, within the United States, unlike many other countries across the world and throughout history, there has never been a coup, nor a serious attempt made by the military to dislodge political control from their duly-constituted elected leaders and to hold it themselves. However, as almost every civil-military relations scholar points out, the absence of coups does not mean that the relationship between civilian and military leaders is healthy, nor absent of conflict altogether. Indeed, a cursory examination of the relationships between Lincoln and McClellan, Truman and MacArthur, and Obama and McChrystal reveals that disagreement between those who serve at the highest levels of civilian government and military organizations is neither rare nor understudied.

What has changed, however, is the degree to which a very specific subset of the American military institution - senior military leaders - have participated in the domestic political process, especially post retirement. Indeed, whereas General Omar Bradley, one of the military's five star generals and a hero of World War Two, famously remarked in 1959, "I am convinced that the best service a retired general can perform is to turn in his tongue along with his suit and mothball his opinions," [@times_bradley_1959], retired generals and admirals today appear frequently in the media, endorse Presidential candidates at higher frequency [@griffiths_not_2019], and frequently serve in government after their retirement from military service [@brooks_dont_2016]. To provide only one such example, following protests across the United States in the summer of 2020 in which President Donald Trump nearly ordered the intervention of active duty military forces into American cities, former Chairman of the Joint Chiefs of Staff, Admiral (Ret) Michael Mullen, was one of at least ten four star retired officers who publicly aired concerns through a variety of mediums about the President's potential use of active-duty military force. ^[Other officers included General (Ret) James Mattis, General (Ret) Mark Kelly, Admiral (Ret) William McRaven, General (Ret) Martin Dempsey, General (Ret) Colin Powell, General (Ret) Michael Hayden, General (Ret) David Petraeus, General (Ret) John Allen, and General (Ret) Vincent Brooks.]. In fact, during an interview in the days which followed the incident, when asked by an NPR reporter why he decided to publicly share his thoughts on the situation, Mullen stated, "we were getting to a point where the country was in jeopardy, which is why I spoke out" [@isaacson_ret_2020; for further analysis, see also @brooks_let_2020].

Mullen's words are incredible from a number of perspectives, and raise important normative questions regarding the military's role in a democratic society [@brooks_let_2020] and the conditions under which the military can and should express dissent to its civilian bosses [@cohn_what_2019]. And while a small but dedicated number of civil-military relations scholars do examine these and related issues quite frequently in today's charged political environment, few, if any, have attempted to systematically examine and explain over time the conditions under which senior military leaders have broken and will break normative traditions and engage in domestic politics. Such an effort would better illumine the understanding of scholars, as well as of the military and society more broadly, concerning why a powerful group of what were once non-political actors have, at least in some cases, suddenly become one.

The central argument made in this dissertation is that increasing political activism at the hands of military officers indicates a shift in the beliefs of the military regarding its central role and purpose in society, or "role beliefs" [@fitch_armed_1998]. This shift, moreover, is the result of two longstanding and interconnected trends: rising political polarization within the domestic political sphere, and the rising prestige and credibility of military institutions afforded by American society. The interaction of these trends explains the remarkable increase in political activity by retired senior military leaders, and portends a shift in the perceived role that these senior military leaders fill vis-a-vis civilian authority - a shift away from professional and obedient "servants" and towards that of "syndicates" or even "saviors" of the American state. This shift, moreover, holds enormous ramifications regarding the organizational culture of the military, its relationship with government and society, and American national security.

# The Field of Civil-Military Relations

My proposed dissertation topic features within the sub-field of civil-military relations, a broad academic sub-field that is generally placed within the International Relations and Comparative Politics sub-field (though within the American context, the sub-field of American Politics is entirely relevant and appropriate). Civil-military relations involve a wide host of significant and important theoretical questions that involve the pull and push of power between institutions, the delineation of authority between military and civilian leaders, and the relationship between a state's military and society.

At its root, the study of civil-military relations examines what Peter Feaver calls the "problematique," or the question of how to ensure that a military is strong enough to defeat threats to the polity without threatening the polity itself [@feaver_civil-military_1996; @feaver_civil-military_1999]. Feaver describes this "problematique" as "vexing" and a great "paradox" in that the military must be "strong enough to prevail in war," yet submissive to authority such that it cannot and will not "destroy or prey on the society it is intended to protect" [@feaver_civil-military_1999, p. 214]. Moreover, the study of civil-military relations often focuses on the relationship between military and government elites, or at what Risa Brooks calls "leadership at the state's apex" [@brooks_risa_integrating_2019, p. 380]. Indeed, it is at this level of interaction between government and military leaders that vital decisions are made, including decisions whether to go to war, how to achieve and implement effective strategy, how to prepare for the next conflict, and how large the military and its budget should be.

A number of scholars have offered solutions to the"problematique" raised by Feaver, but none has been more lasting and consequential than the late Samuel Huntington, whose 1957 seminal work, _The Soldier and the State: The Theory and Politics of Civil-Military Relations_, addresses this fundamental problem. Indeed, even more than 60 years after he wrote his book, Huntington's work often serves as the point of departure for many scholarly works on topics concerning civil-military relations. Although there have been many critiques of Huntington over the years [for example, see @janowitz_professional_1960; @finer_man_1962; @cohen_supreme_2003; @feaver_armed_2003; @brooks_paradoxes_2020], it is absolutely essential to understand Huntington's normative conception of effective civil-military relations, and what lies at the heart of it - a theory of civilian control of the military that maximizes a concept that Huntington called "military professionalism" [@huntington_soldier_1957].

In _The Soldier and the State_, Huntington asserts that two distinct forces would always be in tension with one another in a liberal democracy such as the United States. The first was a "functional imperative," which Huntington said drove the requirement that a military be strong enough and capable of defeating the military threats it faced, and the second was a "societal imperative," which Huntington claims actually constrains the military in that the military can never become so alienated from society such that the military no longer represents the liberal democracy it defends [@huntington_soldier_1957, 1-3]. The societal imperative, Huntington claimed, was a further function of two distinct features of the United States. The first was an "ideological constant," which referred to America's political ideology of liberalism [@huntington_soldier_1957, 143-162]. Huntington argued that liberalism as a political ideology stressed "individualism" [@huntington_soldier_1957, 90], and that American society was, since its founding, suspicious of standing military organizations and professional military forces. The second feature of the societal imperative, however, was the US Constitution, which Huntington claimed to be "conservative," and written by "men who feared concentrated political power and who provided for the widespread dispersion of that power among numerous governmental units" [@huntington_soldier_1957, 143]. Huntington claimed that the tension generated by these two imperatives - one functional stemming from external threats, the other societal stemming from the character of American political ideology and structure - needed to be solved. Writing in the context of the seemingly long and dangerous Cold War, Huntington's point is that America needs strong military that can win on the battlefield, but cautions that such a military cannot fundamentally alienate the character of American liberal society. Both goals - military security and civilian control - had to be obtained.

Figure 1, below, shows a visual diagram of Huntington's logic in a political science framework [@feaver_civil-military_1996]. As Peter Feaver points out in a 1996 article, Huntington did not use such language when developing and explaining his concept. According to Huntington, the ultimate goal, or dependent variable, of civil-military relations in the American context is achieving both military security and maintaining civilian supremacy over the military. Huntington viewed the key independent variable that enables the achievement of these goals while also resolving the tension generated by the opposing functional and societal imperatives as the adoption of a type of civil-military relations framework which he calls, "objective civilian control" of the military [@huntington_soldier_1957, 83]. In such a model of "objective civilian control," Huntington envisioned civilians granting the military a high degree of professional autonomy, which he believed would simultaneously be "most conducive to the emergence of professional attitudes and behavior among members of the officer corps" as well as the best way to minimize the military's political power by rendering the military "politically sterile and neutral" [@huntington_soldier_1957, 83-84], and thus preserving civilian supremacy.

```{r figure one, include=FALSE, echo=FALSE}
huntington_dag <- dagify(
    MSCS ~ CC,
    CC ~ FI + SI,
    SI~I + S,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "MSCS", 1, .5,
      "CC", .7, .5,
      "FI", .5, .6,
      "SI", .5, .4,
      "I", .3, .3,
      "S",.7, .3,
    ))

huntington_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

huntington_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("MSCS"="Military Security and\n Civilian Supremacy", 
                     "CC"="Patterns of\n Civilian Control\n 'Objective Control'", 
                     "FI"="Functional\n Imperative", 
                     "SI"="Societal\n Imperative",
                     "I"="Liberal Ideology",
                     "S"="Constitutional Structure")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 1: Samuel Huntington's Theory Of Civilian Control of the Military", subtitle="As Explained in The Soldier and The State, 1957")

```

Huntington's argument in support of "objective control" of the military stands in contra-distinction, as he explains, to an alternate model of "subjective control" in which civilian leaders limit the power of the military by treating it as any other civilian interest group [@huntington_soldier_1957, 80-83]. In fact, Huntington admits that the US Constitution actually envisioned and established a form of subjective vice objective scheme of civilian control of the military [@huntington_soldier_1957, 163-168]. What the framers did not see, nor could they, argues Huntington, was the rise of the military as a profession in response to ever-increasing and dangerous external threats. It was because of these pernicious threats, argues Huntington, which the US military had witnessed, confronted, and overcome during the 19th and 20th centuries in World War One, World War Two, and as Huntington writes, the Cold War, as well as the destructive capability of modern weapons and warfare in general, that the military profession had emerged [@huntington_soldier_1957]. Huntington claims that Cold-War military officers, unlike the corporate body of officers that existed at the nation's founding, constitute a profession and thus, are part of a unique and distinct organization that differs from other occupations: professions practice and develop a special body of knowledge and expertise; professions exercise a  "social responsibility" by providing a vital service to society; and professions have a degree of "corporateness" in that they develop and formalize standards by which members of the profession are certified to practice [@huntington_soldier_1957, 11-18].

Numerous critiques surrounding the Huntingtonian conception of _military professionalism_ have been levied over the years. Most of these critiques, in one way or another, charge that a strict theoretical and practical separation between military and political operating spheres, which Huntington to a certain degree certainly argued for, is fundamentally flawed. Many, if not all, of these critiques cite Clausewitz's famous dictum that war is a "mere continuation of policy by other means" [@clausewitz_war_1976, 69, 605], and thus point out that by its very nature, war requires the highest integration - not segregation - of political and military activities. Moreover, these critiques often point out, correctly in this author's opinion, that the conduct of war is more complicated than civilians simply telling the military the goals it wishes to achieve, and then allowing the military to determine how best to accomplish these goals. Rather, the development and implementation of strategy, especially during a time of war, requires an "iterative" dialogue between military leaders and civilians to ensure that effective policies are adopted [@rapp_civil-military_2015, 20]. Others argue that Huntington's conception of military professionalism is too narrow to be of practical use in the contemporary period in that such a model actually enables, rather than constrains, inappropriate political activity by military officers [@brooks_paradoxes_2020]. Risa Brooks charges that military officers who strongly identify with an ethic that stresses remaining apolitical at all costs, which she alleges follows from a Huntingtonian conception of military professionalism, may develop "blind spots such that they fail to recognize the political content or impact of their actions" [@brooks_paradoxes_2020, 17]. Whether military leaders recognize it or not, Brooks argues, military leaders comprise a profession that is inherently political, and such a reality ought to be acknowledged and embraced [@brooks_paradoxes_2020].

Regardless of whether and the degree to which one agrees or disagrees with Huntington's concept of military professionalism, a significant number of scholars note that an enduring issue involves attempting to define where the line between appropriate political expression and inappropriate partisan activity should be drawn [@owens_military_2015; @golby_beyond_2015; @brooks_paradoxes_2020]. It is a difficult task for many reasons because, as Mackubin Thomas Owens points out, civil-military relations in the United States can be viewed as a "bargain" [@owens_us_2011] between the American people, the military, and the government that has to account for constantly-changing "political, social, technological, or geopolitical changes" [@owens_is_2016]. In other words, the ground is always moving for the actors involved in civil-military relations.

# Framing the DV: Military Intervention in Politics and Military "Role Beliefs"

Given Huntington's normative model and the critiques his model has been subjected to, it is important to understand what frameworks exist for understanding the debates surrounding why the military would or should intervene in politics, and to what degree. Here again, the literature is vast. Scholars from the field of comparative politics have generally tackled the question of why militaries intervene in political matters, likely as a result of the variation that is available in country-specific or cross-national studies. It is vital to briefly understand a couple of these important frameworks.

The British political scientist, S.E. Finer, argues that military intervention in politics is ultimately a function of both the opportunity and desire to do so [@finer_man_1962]. Finer first identifies several factors that should inhibit the military from intervening in politics, including the concept of military professionalism (as espoused by Huntington), strict adherence to the principle of civilian supremacy, and the military's fear that intervention in politics would lead to it fracturing internally, and thus failing to remain a cohesive fighting force [@finer_man_1962]. Finer delivers a rather stinging critique of Huntingtonian military professionalism, and argues that although professionalism should work to focus the military on itself and thus, away from politics, the danger of the concept of military professionalism is that it can also be abused, particularly by military leaders who adopt the view that they serve the larger concept of "the state" rather than a particular civilian government [@finer_man_1962, 22-23]. Finer argues that military leaders throughout history, including von Seeckt and Gröner under the German Weimar Republic, and MacArthur under Truman in the United States, inappropriately intervened in politics precisely because their conceptions of "professionalism" demanded that they did so [@finer_man_1962], a critique of military professionalism that Risa Brooks shares [@brooks_paradoxes_2020]. Finer's whole point is that the principle of civilian control - not military professionalism - is and ought to serve as the only bulwark to keep the military out of politics.

Finer further describes at least three motives that militaries might have for intervening in politics.  He argues that soldiers might adopt a specific form of "manifest destiny," or a worldview in which the military acts as "saviors" of the country [@finer_man_1962, 28-29]. This motive, Finer claims, arises because of the confluence of the military's unique mission, sense of power, and claim of being "outside party politics" [@finer_man_1962, 29]. A second motive that militaries might possess which disposes them to intervene in politics is what Finer calls the "unique identification with the national interest" [@finer_man_1962, 30]. Finer argues that militaries which act from such a motive might view themselves as the "custodians" of the "national interest" [@finer_man_1962, 30], a role that can take different forms, including a "duty to arbitrate or veto...to exercise it when some convulsion or decision of the civil authorities seems to them [the military] to threaten what they think are the permanent interests of the nation" [@finer_man_1962, 31]. Finally, Finer posits that militaries intervene in politics because of "the sectional interest," a wide-ranging set of reasons based on social class, regional affiliation, personal career motivations, or organizational interests [@finer_man_1962, 34-53].

Finer further classifies military intervention as one of four ascending levels. At the bottom or first level, which he calls "influence," Finer views the military as making appeals "to reason or emotion" in an attempt to persuade the civilian government to adopt a certain course of action [@finer_man_1962, 77]. Finer adds that intervention by the military at this level is not problematic and entirely "consistent with the supremacy of the civil power" [@finer_man_1962, 77]. The next level of intervention Finer labels "pressures or blackmail," describing efforts by the military to "convince the civil power by threat of some sanction" [@finer_man_1962, 77]. Finer adds that military intervention via this level is "on the margin," referring to his belief that while some cases may not be problematic, many other forms, such as the "threat to stage a coup," are normatively wrong [@finer_man_1962, 77]. The third level of intervention, "displacement," refers to the "removal" of government officials by the military as a result of "violence or the threat of violence," and the fourth level, "supplantment," occurs when the military "sweeps away the civilian regime and establishes the military in its place" [@finer_man_1962, 78].

Another helpful conceptualization of military intervention in politics is the concept of military "role beliefs," introduced by J. Samuel Fitch in his book, _The Armed Forces and Democracy in Latin America._ Through rich interview data conducted in the 1990s, Fitch traces how leaders within the militaries of Argentina and Ecuador adopted one of five distinct "role beliefs," which describe the perceptions held by these officers of their normative role in politics [@fitch_armed_1998, 61]. The first type of role belief is "democratic professionalism," and involves both "a subjective commitment to democracy as the preferred political system," and a "belief in a professional ethic that differentiates and separates military and political affairs" [@fitch_armed_1998, 66]. A second "role belief" is "classical professionalism," and differs from the first in that while it "argues for the subordination of the military, in both political and policy terms, to the constitutionally designated civilian authorities", it does not espouse a "normative commitment to a democratic regime" [@fitch_armed_1998, 66-67]. A third type of "role belief" that military offices can adopt is that of a "power factor," or a belief that within the regime, "the military should have some degree of influence in deference to its vital societal role" [@fitch_armed_1998, 67]. The fourth type of "role belief" militaries can adopt is one of "conditional separation" to civilian leaders, and refers to the notion that while the military should stay out of politics "in normal circumstances", the advent of "crisis situations" can and does warrant the military to "safeguard the highest interests of the nation," and, as Fitch adds, "even if that entails overthrowing the incumbent government and assuming power temporarily" [@fitch_armed_1998, 68]. A final possible role belief militaries can adopt is a "tutelary" model in which the civilian government grants the military both a "high degree of autonomy" and the ability to become a regular and "institutional voice in policy matters" [@fitch_armed_1998, 68-69]. According to this "role belief," the military believes itself to be the final "guardian of national security," and thus that it has "the right" to participate in a wide variety of policy debates that impact the state [@fitch_armed_1998, 69]. Fitch argues that the particular "role beliefs" militaries adopt are a function of three primary factors: the lessons militaries draw from previous domestic intervention in politics, a wide-ranging assessment of current domestic indicators (the economy, the effectiveness of institutions, etc.), and the prevailing  international context [@fitch_armed_1998].

Within the US context, scholars have long been interested in the specific political activities that the military engages in. These studies have tended to examine specific political behaviors such as voting [@dempsey_our_2010], social media use [@urben_like_2017], political endorsements [@griffiths_not_2019], and attitudes towards political engagement [@urben_party_2013; @urben_wearing_2014], but they have not attempted to systematically explore these changes over time. Additionally, while Risa Brooks has created a helpful typology of political activities the US military engages in, ranging from "public appeals" that engage the public, to "shoulder tapping" focused exclusively on Congress [@brooks_militaries_2009], her typology also does not explain variation in these activities over time. While the domestic environment in Latin America is vastly different than that in the United States, Fitch's work underscores and points to the importance of civilian and military leaders committing themselves to the supremacy of both democratic institutions and democratic rule [@fitch_armed_1998]. And while there has arguably been more variation in Latin American militaries intervening in politics, I believe that there is sufficient evidence that indicates a subtle but important shift in the US military's beliefs regarding its central role and purpose has occurred, and that a systematic examination of this shift is warranted.

# Framing the IVs: Political Polarization and Military Prestige

## Political Polarization

Within the American Politics literature, there is broad recognition that political polarization levels in contemporary America are higher now than at any point since the end of the American Civil War [@hare_polarization_2014], and there are multiple explanations for what has caused this trend, ranging from a growing divide over matters pertaining to race and religion [@abramowitz_great_2018], the size and role of government [@sides_two_2015], and disagreements over specific policy issues [@mann_its_2012]. Despite the wide-ranging scholarship within the field of American Politics that explores the facets of political polarization, virtually all scholars agree that contemporary levels of political polarization pose myriad problems.

Scholars of international relations have, more recently, examined the pernicious impacts of polarization on foreign policy. Kenneth Schultz, for instance, underscored four challenges that have emerged as a result of America's polarized domestic politics, including the ability to obtain bipartisan support for "risky undertakings," the ability to learn the "lessons of foreign policy failure," the inability to "make long term commitments to allies and adversaries" as a result of "dramatic policy swings from one administration to another," and the "vulnerability" of America's electoral system to foreign third party adversaries [@schultz_perils_2017]. Because the military is a central actor in shaping America's foreign policy, obtaining a better understanding of how political polarization shapes various military institutions will yield findings that may help policymakers and academics help solve the problems Schultz identifies.

But a more central concern of mine is understanding how and why political polarization might impact how the military views its central role in American politics. I argue that there are at least three ways through which increasing polarization can shape the apolitical and non-partisan ethos of the military, and thus, can alter how the military views its role in politics. The first is that there is increasing difficulty for military leaders to remain politically neutral, non-partisan, or otherwise apolitical - a hallmark of Huntingtonian military professionalism. Fewer and fewer issue areas and policy matters are outside of the "spell" of polarization, owing to a phenomenon scholars of American politics refer to as "conflict extension" [@layman_party_2002]. In previous eras of polarization, whereas a single issue may have served as the "battleground" for partisan conflict, this is no longer true today, as research has consistently shown that voters who identify as liberals are more liberal across multiple issue categories and that the same trend holds for conservatives [@abramowitz_great_2018]. As a result, whereas the prospect for cooperation on issues that were not politically polarized remained high in the past, today the prospect for cooperation in general is far lower [@layman_activists_2010].

A second challenge posed by rising polarization stems from the drastic increase of negative feelings toward those on the other side of the political spectrum, a phenomenon known as affective polarization or "negative partisanship"[@abramowitz_great_2018]. Just as contemporary polarization is marked by citizens who hold increasingly strong and negative feelings towards those of a different political stripe, there is likely a tendency for members of the military institution to harbor similar feelings and attitudes. One might object, pointing to the training and professional development military leaders receive that emphasizes the importance of upholding an apolitical and non-partisan ethic, but even a cursory review of the evidence (members of the military increasingly making presidential endorsements [@griffiths_not_2019] and posting political messages on social media [@urben_like_2017] ) suggests this is not the case.

A third, and I contend the most impactful, challenge posed by polarization to the normative conceptions of the military's role in politics stems from the fact that our polarized climate is marked by the belief among individuals that core issues - such as those that involve attitudes about race, religion, and multiple layers of not just individual, but also national identity - are at stake in the current polarized political climate. As the rise of "identity politics" illustrates, some of the most polarizing issues today are those that are related to an individual's identity, such as race, gender, and immigration [@sides_identity_2018]. And at the national level, the same phenomenon holds: as Vice President Joe Biden and President Donald Trump wind down the final days of their presidential campaigns, both claim that "the soul of the nation" is at stake [@dias_biden_2020].

## Military Prestige

With respect to the military institution itself, one of the most massive and influential changes that has occurred over the past roughly seventy years is the increase in overall prestige, influence, and public credibility of the military institution. While key legislative acts passed since World War Two, to include the National Security Act of 1947 and the Goldwater-Nichols Act of 1946, may have fixed important gaps in America's ability to meet its national security requirements [@inboden_national_2017], they also cemented the military as one of the dominant institutions and actors within American politics. Over the same time period, while there have been spikes as a result of exogenous events such as wars and scandals that have involved the military, public opinion towards the military has increased. Consider that in the late 1970s, as a likely result of public opinion over the Vietnam War, 54% of Americans indicated that they trusted the military "a great deal or quite a lot," whereas in 2019, 73% of Americas expressed the same sentiments [@gallup_polls_confidence_2007]. Compared to other government institutions, today the American public trusts the military to a remarkably high degree. According to a recent Pew Research Center poll, 80% of the American public expects the military to act in the best interests of the American public, compared to business leaders (45%), the media (40%), and elected officials (25%) [@johnson_trust_2018].

Coinciding with these developments, scholars began exploring whether there was an unhealthy "gap" forming between American society and the military, especially after the US ended conscription and formed the all-volunteer force in 1973 [@feaver_soldiers_2001; @schake_warriors_2016]. While there is no definitive answer as to whether a broad "gap" exists, scholars have noted the existence of several smaller "gaps," including a "partisan gap" [@golby_duty_2011] between an officer corps that strongly (though not exclusively) identifies with the Republican party and potential Democratic presidential administrations, and a "familiarity gap" between an all-volunteer military that has endured repeated deployments since September 11, 2001 and American society, especially younger Americans [@golby_thanks_2016; @colford_young_2016]. A major concern of scholars is that to the degree that these gaps exist, these potentially facilitate the formation of unhealthy attitudes regarding the superiority of military service over and above other types of public service, and further reinforces the military's isolation from society [@fallows_tragedy_2014] while also negatively impacting democracy [@friend_military_2018]. According to Mara Karlin and Alice Friend, the act of "military worship" - esteeming the military more than society ought to - can, over the long haul, be problematic for civilian control of the military [@friend_military_2018].

We further know, as scholars have argued, that when elites possess a high degree of credibility, they possess the ability to influence and shape public opinion. In his influential study, _The Nature and Origins of Mass Opinion_, John Zaller argues that formations in mass public opinion are the result of three interacting components: "variation in the information carried in elite discourse, individual differences in attention to this information, and individual differences in political values and other predispositions" [@zaller_nature_1992]. Zaller argues that throughout modern America, most citizens, as non-elites, must rely heavily on the messages that come from elites regarding various topics and issue areas [@zaller_nature_1992]. Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send, in order to be understood by regular citizens who are part of the public, must be "short, simple, and highly thematic" [@zaller_nature_1992]. In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information, but rather acknowledging that most citizens care most about a world that is not "beyond their personal experience" [@zaller_nature_1992]. Thus, his main point is simply that elites play a significant role in forming and swaying public opinion.

Therefore, as the public's trust in the military has grown, as gaps between segments of society and the all-volunteer force have emerged, and as military service has come to be viewed as somehow superior to other forms of civilian service, the latent power and influence of the military has grown significantly. This is especially true of military elites, who hold substantial positions of authority and responsibility. However, an increase in the power and influence of the military institution may not indicate that the military will actually use it in particularly harmful ways, but merely that it has the potential to do so, as the American people tend to listen to those whom they trust and revere.

# My Puzzle, Where it Fits, and Why It Matters

The central question I wish to explore in my dissertation is whether and why the US military, or portions of it, have become increasingly politically active, and whether such a shift portends a shift in the military's own "role beliefs" concerning how it should and does relate to American government and society. As such, my dissertation will center on fundamental and longstanding debates about the proper role of the military with respect to government. I wish to explore changes in these concepts over time and in historical perspective within the United States.

While there have been several rich and important studies that explore the attitudes of military officers towards various forms of political behavior [@dempsey_our_2010; @urben_party_2013; @urben_wearing_2014], these studies have not primarily sought to explain why these attitudes have changed over time. Similarly, while there have been a number of important studies exploring the impacts of political polarization on the credibility society places in the military as an institution [@robinson_michael_danger_2018; @robinson_who_2019], the dependent variable in these studies was the credibility that society places in the military. There have been a number of impressionistic pieces noting the potentially detrimental impacts of polarization and politicization on the military institution [@feaver_we_2016; @barno_how_2016; @golby_jim_keep_2020; @reid_retired_2020], but these works, too, have not sought to empirically explain the impacts of these changes over time. It is my opinion that civil-military relations scholars need grounding in theory, specific to the context of the United States, that seeks to explore how rising political polarization and military influence interact, and that such a theory needs to be supported by rigorous empirical work. I wish to close this gap, as doing so will enable scholars and practitioners, to include members of the military, to understand the answers to some very important questions, including:

* Has the answer to the question, "what does it mean to be a military professional?" changed over the past several decades, and if so, why?
* Among whom within the military institution is a shift in normative "role beliefs" occurring, and why?
* When did such a shift start, and have there been similar periods in American history?
* If there is a shift among portions of the military in "role beliefs," how does such a shift impact the military's organizational culture?
* Are some of the most contemporary polarizing domestic issues compatible with historically normative conceptions of military professionalism, such as the principles of civilian supremacy and a non-partisan ethos? Why or why not?
* What are the key challenges or obstacles to the military regaining and sustaining an effective non-partisan ethos? 
* Should we expect these challenges and obstacles to endure?

Answers to these questions will be of particular importance to furthering our understanding of how the US military has responded to epochs of polarizing politics before, and how changes in levels of institutional prestige impact these responses. I suspect that my dissertation will show that any contemporary shift in military role beliefs is not the first shift that has occurred in American history, that such a shift has long been underway, and that particular aspects of today's polarized climate portend that the military will wrestle with the issue of shifting role beliefs for years to come. For these reasons, and more, a systematic investigation of the impact of polarizing politics and institutional prestige on the US military's beliefs concerning its role in government should be undertaken.

# The Analytical Framework - Initial Argument

## The Impact of Polarization and Military Prestige on US Military Role Beliefs

In this section of the prospectus, I will lay out an initial argument which I intend to explore and test in my dissertation. Figure 2, below, shows an initial conception of the linkages between the variables discussed in the previous section of the prospectus. From left to right, the model suggests that political polarization, prestige and credibility, and war (independent variables) all shape how the military views its fundamental role and purpose within the government and society (the dependent variable).

Borrowing from Fitch, I will use the term military "role beliefs," but I will alter the concept he sets forth ever so slightly. Whereas he used the term broadly to describe the set of norms, beliefs, and attitudes that the military holds regarding its "role in politics" [@fitch_armed_1998], I argue that we can be more specific in the US context. In particular, the "role beliefs" of the military comprise the answer to the question, "what does it mean to be a military professional?"

After surveying the rich normative literature on the topic of military professionalism, to include critiques of the concept, there are, I argue, two central tenets that stand out above others when asking the question, "what does it mean to be a military professional?" These tenets are adherence to the principle of civilian control of the military, and confinement of one's knowledge and skills to relevant military expertise. The first principle of course deals with who is in charge of the military, and the second deals with for what purpose the military exists. The degree to which the military deviates from one or both of these central tenets indicates a shift in the role beliefs the military holds. Figure Two, below, is a visual illustration of the variables of interest in this study. They include the dependent variable, military role beliefs, and the independent variables of war, political polarization, and prestige/credibility of the military.

```{r figure two, include=FALSE, echo=FALSE}
problem_dag <- dagify(
    RB ~ PC + PP + W,
    RB~PP,
    PC ~ PP,
    coords = tribble(
      ~ name,      ~ x,    ~ y,
      "RB", 1, .5,
      "PC", .6, .5,
      "PP", .3, .7,
      "W", .8,.7,
    ))

problem_dag %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text_repel(aes(label=name), show.legend=FALSE)+
  theme_dag()
```

```{r include=TRUE, echo=FALSE, fig.height=5, fig.width=7}

problem_dag %>% tidy_dagitty() %>% 
  dag_label(labels=c("RB"="Military Role Beliefs", 
                     "PC"="Prestige and Credibility", 
                     "PP"="Political Polarization",
                     "W"="War")) %>% 
  ggplot(aes(x=x, y=y, xend=xend, yend=yend))+
  geom_dag_edges()+
  geom_dag_point()+
  geom_dag_text()+ 
  geom_dag_label_repel(aes(label=label, fill=label), col="white", show.legend=FALSE)+ theme_dag() + labs(title="Figure 2", subtitle="Determinants of US Military Role Beliefs")

```

Beginning with the impact of war on military role beliefs, I have chosen to draw an arrow directly from war to role beliefs, and not from war to political polarization, or vice versa. There is sound theoretical expectation that while the military is actively engaged in fighting a conflict, they will focus on prevailing in that conflict, and thus, that active combat has a dampening effect on the military's propensity to inappropriately intervene in politics. Such an argument follows a similar (though not identical) argument made by Michael Desch, who theorized that civilians would be in a better position to control the military when external threat levels are high [@desch_civilian_2001].

War, vice threat, is shown in this model, however, following recent scholarship that argues that threats in themselves hold little if any unifying power, and thus, do not attenuate polarization levels [@myrick_external_2020]. Other scholars have argued that combat operations and war tend to focus the military inward. Don Snider, for instance, argues that military institutions tend to diverge away from their professional nature as war fighters and towards that of a bureaucracy as combat operations wind down and as peacetime ushers in changes such as force reductions and a return to training for combat [@snider_once_2012]. I follow Snider's expectation and argument, and conclude that war is the appropriate variable to show here.

In terms of the two central tenets of the professional nature of the military, I argue that war can theoretically impact both, but its main impact is on expertise. Winning in combat demands that the military develop and implement unique military expertise to prevail in war. Conversely, winning in combat does not mean that the military will necessarily adhere to the principle of civilian control, as American history aptly demonstrates (e.g., Truman and MacArthur), but the presence of active combat does predict that military intervention in politics during wartime at the hands of the military will likely center on the conduct of the war itself, and not on domestic political matters that fall outside of military expertise.

The next variable to theoretically unpack is political polarization, and admittedly, this is a difficult endeavor. However, previous literature in American politics enables the development of two plausible relationships and corresponding mechanisms, each stemming from a particular conception or definition of polarization. If one conceives of and focuses on polarization as the ideological widening mainly between political elites, such as members of Congress [@hare_polarization_2014], then one possible mechanism by which the military intervenes in politics is by adopting a belief that elites from both dominant political parties have failed to advance and implement an effective governing agenda. Thus, the military steps in to do what political elites could not do. Scholars of comparative politics have noted such a dynamic occurring among militaries in other parts of the world [@fitch_armed_1998].

A second, and I believe the more likely, relationship between political polarization and military role beliefs, particularly in the contemporary United States, results from a conception of polarization as the strength of feelings concerning a number of ideological issues. This is a view of polarization that is consistent with the concept of affective polarization, or "negative partisanship" [@abramowitz_great_2018]. If the variables of polarization and military role beliefs are linked in this way, the military would not intervene as a monolithic block due to a frustration with the political class's inability writ large to effectively govern. Rather, certain military elites, who share beliefs with a polarized electorate on any number of ideological issues, such as race, gender, and immigration [@abramowitz_great_2018], will selectively intervene in order to advance these very beliefs. In other words, this mechanism posits that certain military elites feel as strongly as average citizens about certain issues, and that they will break from historical apolitical or non-partisan military traditions and norms in order to advance these causes.

This conception adopts a view such that despite the military organization's emphasis on apolitical traditions, beliefs, and norms, the forces of political polarization and the ideologies that underlie it have the potential to prove stronger. Additionally, political polarization may impact both of the tenets of military professionalism (adherence to the principle of civilian control, and adherence to relevant military expertise). High polarization may impact the principle of civilian control because specific military leaders may hold differing ideological beliefs than their civilian bosses, and it may impact expertise because certain military leaders may adopt a view that they can perform the functions of their civilian leaders better, even if these functions fall well outside the confines of what traditionally constitutes military expertise.

In short, military leaders are experiencing and reacting to the very same phenomena that scholars of American politics have identified as impacting American society more broadly: the phenomena of "conflict extension" [@layman_party_2002] and "negative partisanship" [@abramowitz_great_2018], and the fact that polarizing domestic disputes involve so many issues that individuals consider a "core" interest or value [@white_values_2002; @mann_its_2012; @sides_identity_2018]. As a result of these dynamics, political polarization has provided the _"motive"_ for certain military leaders to intervene in domestic politics, which is exactly what Finer warned us about 58 years ago [@finer_man_1962]. When polarization is not high, there simply is less of a motive for the military, to include military elites, to deviate from their historical non-partisan and apolitical traditions.

\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{How Features of Contemporary Political Polarization Impact Military Role Beliefs} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    \textbf{Features of Increasing Political Polarization} & \textbf{Associated Mechanism} & \textbf{Potential Impact on Military Role Beliefs}  \\ \hline
    \textbf{Conflict Extension} & Less politically-neutral space exists as partisan disagreements subsume a greater number/scope of issues. & Military Professionals and the Institution find it increasingly difficult to "stay out" of partisan politics as the territory of what is considered partisan politics expands. \\ \hline
    \textbf{Negative Partisanship} & Strong feelings of personal dislike and disdain arise against those who hold to opposing political views. There is a belief that those on the other side of the political aisle are not just mistaken over policy, but greatly wrong and flawed as individuals. & Even among military leaders, there is a greater willingness to confront or take issue with those who hold opposing views. Like average citizens, military elites feel they should speak up and against those whom they feel hold threatening political views.\\ \hline
    \textbf{Individuals believe that "core" issues are increasingly at stake} & The salience of one's identity and core values is raised as partisans believe essential issues are at stake in the political arena. Associated with this is a belief that not just policy preference, but moral right and wrong, are at stake. & Military professionals face an increasingly difficult task of balancing the multiple loyalties they owe, especially as the objects of these loyalties are increasingly associated with specific ideological or partisan causes, and are attached to notions of moral right and wrong. \\ \hline
\end{tabular}
\label{table:comparison} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

Figure Two also argues that the level of prestige and credibility afforded to the military also impact military role beliefs. Here, I view prestige and credibility in both an _institutional_ or _organizational_ sense, as well as an _individualistic_ sense. It is important to distinguish the two because of the variation that can occur within each. For example, General (Ret.) David Petraeus enjoyed an outsized level of personal prestige and credibility during the height of the surge in Iraq, whereas Lieutenant General (Ret.) Ricardo Sanchez likely lost, at least temporarily, some individual prestige and credibility after being removed from commanding US forces during the first year of Operation Iraqi Freedom, and as allegations of detainee abuse at the hands of US forces at the abu-Ghraib prison, near Baghdad, surfaced [@freed_last_2011]. Both officers, however, served in their positions at a time when the US military commanded a relatively high level of institutional prestige and credibility relative to other US institutions [@gallup_polls_confidence_2020].

Whereas high political polarization provides the "motive" to intervene, I contend that high levels of prestige and credibility provide the other missing link - the _"opportunity"_ - to do so [@finer_man_1962]. High levels of prestige and credibility, often correlated with high experience levels in a specialized area, serve as latent power to influence a number of important audiences, to include the general public  [@zaller_nature_1992], key government figures [@saunders_no_2017], and other organizations and institutions. Zaller argues that most citizens, as non-elites, must rely heavily on the messages that come from elites regarding complex topics and issue areas [@zaller_nature_1992]. Zaller further argues that the messages that elites send is necessarily "stereotyped" in some way: that is, the messages that elites send must be "short, simple, and highly thematic" [@zaller_nature_1992]. In developing this argument, Zaller is not claiming that regular citizens are uninformed or careless people who are incapable of processing complex information, but rather acknowledging that most citizens care most about a world that is not "beyond their personal experience" [@zaller_nature_1992]. Thus, his main point is simply that elites play a significant role in forming and swaying public opinion. Therefore, when the military institution or when military elites command a high degree of prestige and credibility, they are in a position to heavily influence those around them. Therefore, high prestige and credibility can, like political polarization, impact both the military's adherence to the principle of civilian control and its willingness to stick to relevant military expertise.

What is less straightforward, however, is unpacking how political polarization impacts the levels of prestige and credibility of the military institution, as Figure Two illustrates. While it is theoretically possible that as polarization rises, the public views the military more favorably relative to other institutions precisely because of its non-partisan ethic and traditions, recent scholarship has shown that the public views and responds to specific military actions almost entirely along partisan lines [@golby_listening_2013; @golby_jim_elite_2017; @robinson_michael_danger_2018]. This has further complicated the military's role in politics, as it essentially means that one's partisan identity heavily conditions one's opinion of the military, and in particular, any incident that involves the military's politicization: if the military or a particular military leader says or does something that aligns with one's ideological views, that particular individual is likely to view the military favorably, while the opposite is also true [@obinson_michael_danger_2018].

This dynamic may impact military role beliefs in important ways. The military has long cherished its ability to hold the public's trust. In fact, former Army Chief of Staff, General (Ret) Ray Odierno used to say, "trust is the bedrock of our profession" @us_army_repairing_2014]. But if the degree of prestige and credibility that the military commands splinters along partisan or ideological lines, the military may believe that it should or must act to repair that gap in trust. Yet doing so, especially along partisan or ideological lines, could cause even more divisions with opposing partisan groups, and the spiral in credibility among segments of the public may continue. Therefore, it is reasonable to conclude that while polarization may impact military prestige and credibility, it is less clear to conclude precisely how so.

When the levels of military prestige and credibility are high, the American public approves of the military's performance and behavior, and thus, there is somewhat of a demand signal for the military to continue doing whatever it has been doing. In the contemporary environment, this might equate to the military continuing to interact or dialogue with the public in a number of ways, such as military participation in public ceremonies, parades, or other high visibility events, or from an individual perspective, though the press, social media, or other outlets. When military credibility and prestige are low, there is less of a demand signal from the public to the military institution or to particular military elites, and thus, there will be fewer interactions between the military or military elites and the public, and thus, fewer opportunities for the military to extend their influence.

## When Polarization and Prestige Collide

With the view that rising political polarization in the US may provide the "motive" and that increasing prestige and credibility may offer the "opportunity" for military intervention in politics [@finer_man_1962], it is now possible to determine how the presence of these factors combine to actually manifest in the central role beliefs that militaries adopt. Ultimately, four possible worlds exist (see Table 2, below). The first is when political polarization and military prestige are both relatively low. When these conditions hold, there is little if any expectation that the military will deviate from its non-partisan ethos. Because polarization is low, there are few if any ideological issues or concerns that may serve as a motive for military intervention. Further, owing to a low degree of prestige and credibility, the military does not have a high degree of latent power, and thus, its ability to have an outsized influence on the public, government figures, and other organizations and institutions, is minimal.

\singlespacing
\begin{table}[ht]
\begin{footnotesize}
\caption{US Military Role Beliefs: A Function of Institutional Prestige and Political Polarization} % title of Table
\centering % used for centering table
\begin{tabular}{|p{5cm}|p{5cm}|p{5cm}|p{5cm}|}
    \hline
    Influence of Variables & \textbf{Low Prestige / Credibility} & \textbf{High Prestige / Credibility}  \\ \hline
    \textbf{Low Political Polarization} & Unlikely deviation from an apolitical and non-partisan ethos. Traditional notions of military professionalism are expected by the military, civilian Leaders, and the public. The military does not hold outsized power to influence the public or those within government, and there are few if any issues that motivate the military to deviate from its traditional roles & Little deviation from an apolitical and non-partisan ethos. The public respects the military institution and there is a demand for military voice, but there are few polarizing issues in domestic politics. The military may weigh in on matters, but it will focus on national security issues that directly impact the military. There is little incentive for the military to take visible stances on domestic topics that do not impact the military directly. \\ \hline
    \textbf{High Political Polarization} & Moderate deviation from an apolitical and non-partisan ethos. Polarization provides a motive for the military to intervene in domestic politics, but the military is constrained from holding an outsized influence in debates on these matters due to its low institutional prestige and credibility &  Highly Likely Deviation from an apolitical and non-partisan ethos. Polarizing issues provide a motive for the military or certain elites to intervene, and high levels of prestige and credibility make such intervention possible\\ \hline
\end{tabular}
\label{table:comparison1} % is used to refer this table in the text
\end{footnotesize}
\end{table}
\doublespacing

The second possible set of conditions is when political polarization is low and institutional prestige is high. When these set of conditions exist, we can expect little deviation by the military from its traditionally non-partisan ethos. The military has the potential to wield an outsized influence on the public and within the government, but because the degree of political polarization is low, there is little opportunity nor incentive for the military to adopt or denounce an issue associated with any partisan or ideological cause. When these conditions are present, the military may be active, but likely on matters that directly relate to American national security.

The third general set of conditions occurs when political polarization is high, but institutional credibility of the military is low. Under these conditions, there is a moderate expectation that the military will deviate from its non-partisan and apolitical traditions. The military does not command the respect of the American public to the degree that it does when the prestige of the institution is high, but because polarization levels are high under these set of conditions, certain military leaders may be motivated to intervene in politics to advance certain ideological or partisan causes. There may be advancement of these causes by certain military elites, but such activities are not likely to feature prominently in the public eye. This may involve more "behind the scenes" activities by particular members of the military or military elites.

The final set of conditions exists when political polarization and institutional prestige and credibility are both high. Under such conditions, the military is highly likely to deviate from its apolitical and non-partisan norms, and thus, to intervene in politics. The high degree of prestige the military institution enjoys means that it possesses the ability to influence a number of important audiences by its words and actions, and the high levels of polarization indicate that there are motives to act. It is under this dangerous set of conditions, I argue, that the American military finds itself in now in 2020, as it did at the end of the American Civil War.

# Hypotheses, Observables, and Plan of the Dissertation

In this section of the prospectus, I will elaborate on what a political scientist would expect to see given the theory advanced in the preceding pages, and describe the overall plan for my dissertation. In short, I plan to implement a mixed-methods approach that incorporates quantitative and qualitative analysis, as well as data from an original survey experiment, over the course of what I have now planned to take seven chapters. My dissertation will hinge on the examination of several hypotheses which must be examined, tested, and evaluated, as doing so will enable me to strengthen the overall argument I wish to make.

In broad terms, I need to look for evidence that will confirm or, alternately, disprove, my theory. I have made the central claim that changes in the way a professional military involves itself in politics, and thus, changes in its central role beliefs, will challenge one or both of two central guiding principles: the principle of civilian control of the military, and confinement of knowledge and advice to the realm of relevant military expertise. The theory advanced thus far further argues that although the presence of active war limits the military's deviation from professional role beliefs, two other domestic factors, political polarization and military prestige/credibility, provide what Finer previously identified as the "motive" and "opportunity" for the military's intervention in politics [@finer_man_1962]. Therefore, I need to show that political polarization, and not another variable or set of variables, truly provides sufficient incentive for military leaders to deviate from non-partisan or apolitical traditions. I also need to show that it is the level of military prestige and credibility which affords military elites the opportunity to engage in politics inappropriately, and not some other variable. As a scope condition, I also need to find evidence that the presence of war truly does limit or constrain the military's ability or desire to intervene inappropriately in politics by the way that I have, building on other scholars, argued: that active war keeps the military focused on predominantly military expertise.

## Chapter 1-2 : Introduction and Theory

In the first chapter, I will frame my research question, lay out my research agenda, and review the relevant literature for my project. This will involve synthesizing several of the key aspects and debates within the field of civil-military relations, much in the same way I have done in the first part of this dissertation prospectus. I will focus my attention on orienting the reader to debates regarding a Huntingtonian conception of military professionalism. I will emphasize the long-running nature of this debate going all the way back to when Huntington first published his book, but point to some of the prominent contemporary critiques regarding his theory [@feaver_civil-military_1996-1; @rapp_civil-military_2015; @golby_why_2017; @karlin_case_2020; @brooks_paradoxes_2020]. In chapter one, I will also introduce the growth of political polarization and military prestige in the latter half of the twentieth century in order to motivate the fundamental question of my dissertation.

In the second chapter, I will introduce, develop, and explain a theory that links increasing political polarization and military prestige to the role beliefs that militaries adopt. Much in the same way that I have summarized in the theory section of this prospectus, I will emphasize that I am building off of the theories advanced by previous scholars, such as Finer and Fitch, and applying these theories to the 21st century US military. Using Table Two as a framework, I will argue that to determine whether the military will stray from its traditional non-partisan and apolitical traditions, the key variables to consider are war, polarization, and prestige, and that the key observables are how the military thinks and behaves about two central tenets: the principle of civilian control of the military, and adherence to relevant military expertise. When war is present, the military will likely focus on winning that war, and developing and focusing on the military expertise to do so. When high levels of polarization and prestige are present, however, there is a strong likelihood that certain military actors may use the credibility they enjoy to advance certain ideological causes, potentially bucking both the principle of civilian control and straying outside of relevant military expertise.

## Chapters 3-5: The Heart of the Empirical Analysis

Chapters 3-5 will comprise the core of the empirical analysis of my dissertation. In Chapter 3, I plan to examine variation in opinion-editorial pieces published by retired military generals and admirals in major newspapers since the end of the Vietnam War. The hypotheses that I will examine and test in this chapter are as follows:

_H1: When political polarization is high, the military (or certain military figures) will increasingly criticize civilian leadership (indicating a drop in the principle of civilian control)._

_H2: When political polarization is low, the military (or certain military figures) will publicly weigh in on traditional national security debates (and not on domestic political concerns or issues outside of relevant military expertise)._

_H3: During periods of high prestige and credibility, the frequency of military involvement in quasi-political activities (such as writing op-eds) will increase relative to periods of low prestige and credibility._

_H4: When both political polarization and high prestige are present, the military will be increasingly vocal and at higher frequency about certain domestic ideological or partisan causes than when these factors are not present._

_H5: During periods of active ground combat, military elites will likely comment on the war, but with less frequency on other domestic issues._

To test these hypotheses, I plan to conduct a quantitative analysis of more than 300 opinion-editorials published by retired military officers in the New York Times, Wall Street Journal, and Washington Post since the early 1970s. I plan to hand code these op-eds and classify each according to a categorical subject type (such as whether the piece covers a topic involving foreign policy, war fighting, military budgets, institutional military culture, social policy (such as who serves in the military), or domestic policy (not directly related to the military), and to also code for several dichotomous variables, such as whether the op-ed sharply criticizes a partisan figure (President or other named official), whether the op-ed clearly adopts a partisan stance on an issue, and whether the op-ed strongly calls for definitive legislative or political activity (such as Congress funding an initiative or signing a treaty).

The dependent variables in my analysis will include the frequency of certain subject types of op-eds (such as domestic policy) and the frequency of the dichotomous variables outlined above. Additionally, I will include controls for the level of polarization present in Congress (using DW-NOMINATE scores) in each year of my analysis as a proxy for the intensity of polarization present in society, and the level of military prestige (measured by Gallup public opinion polling that measures the level of trust in the military institution by American society)^[I may also use alternate measurements, such as measures of affective polarization, which will involve measuring the difference in presidential approval ratings between Republicans and Democrats for each year of my analysis.] Because the data covers the past 45 years, from the 1970s to the present day, the data allow for one to examine variation of a quasi-political behavior by the military (writing op-eds) while holding the sources constant across a wide period of time. Moreover, the period of time examined allows the researcher to observe the impacts of exogenous shocks on such behavior (such as the presence of wars, conflict, etc.). Moreover, over the same time period, both of my independent variables of interest, polarization and prestige of the military, have varied considerably, enabling for a fairly clear measurement strategy.

What the data and the ensuing analysis from Chapter 3 cannot answer, however, is the degree to which variation in the observed military behavior is a function of changes in the editorial sources themselves over the same time period. In other words, it will be difficult to determine the degree to which the New York Times, for instance, has shifted to the political left over the past five decades (if any), and what impact these changes may have in determining the content type of its editorials. As far as connecting these initial five hypotheses to my overall theory and argument, I do not think that one hypothesis serves as more critical than any of the others, but rather that together, all five of the hypotheses lend convincing support for my theory. Writing an op-ed is a uniquely quasi-political activity in the sense that an op-ed can be very political or, alternately, entirely apolitical, depending on the content of the op-ed. But unlike tweets, facebook posts, are other modern forms of military elite speech that scholars have analyzed [@robinson_michael_danger_2018], op-eds must be deliberately written, and contain an overall argument. Publishing an op-ed in a major newspaper is also a somewhat competitive process, indicating that the author likely desired or greatly desired to make his or her opinions known. By analyzing the content of op-eds authored by retired flag officers over time and by controlling for the variables outlined above, I will be able to detect the degree to which there have been changes in the willingness of military elites to openly criticize civilian leaders and to weigh in on matters outside of their traditional military expertise, which are both critical measures of the military's role beliefs.

In Chapter 4, I plan to conduct a mixed-methods analysis of the beliefs of senior military officers regarding the role of the military in politics in historical perspective. In order to do this, I plan to examine the behavior of individual high ranking officers - in fact, the highest ranking officers - Chairmen of the Joint Chiefs of Staff, and military service chiefs (such as the Chief of Staff of the Army, the Chief of Naval Operations, the Commandant of the Marine Corps, the Chief of Staff of the Air Force, and the Commandant of the Coast Guard) since World War Two. Examining the behavior of these leaders (roughly 120 in total), particularly during their tenure in the position as head of their respective services (or the military in the case of the Chairman of the Joint Chiefs of Staff) and in each's retirement years, will shed light on what each leader thought about the military's role in politics, and whether each officer lived out such beliefs in their retirement years.

Focusing on the highest ranking military officers is especially important, as it enables a researcher to examine if there have been changes in what these leaders - as the "stewards of the profession" [@us_department_of_the_army_adp_2019, p. 1-9] - have thought regarding the military's role in politics.  Additionally, focusing on the individual enables the researcher to control for individual-level impacts, such as shocks to a specific officer's level of prestige or credibility, that may impact the dependent variable, individual acts of political behavior by military elites. The hypotheses that I wish to examine and test in this chapter include:

_H6: Officers who endure a decline in either their individual or institutional level of prestige will be less politically active than those who experience either a relative boost or no change._

_H7: When polarization and military prestige are low, senior military elites are less likely to weigh in on turbulent domestic political events than when polarization and military prestige are high._

_H8: As polarization and prestige increase, there will be a gap between the "role beliefs" espoused by senior military leaders while on active duty and the "role beliefs" these same leaders live out during retirement._

The dependent variables in this chapter will include political acts of behavior or seemingly political behavior, to include: writing opinion editorials (building off of data in Chapter 3), providing congressional testimony (on active duty as well as during retirement), and weighing in on turbulent political events in the form of speeches, articles, commentary, etc.^[I would include other actions, such as endorsing a Presidential candidate for office or speaking at a political convention, but these more "obvious" forms of political behavior did not start until at the late 1980s (publicly endorsing a presidential candidate), and even later (for speaking at a national political convention). Thus, I need to account for this type of "discontinuity" in my results, i.e. somewhere in the late 1980s, military officers began these particular types of behavior in earnest.] I will also record the occupation or primary role that each officer held in retirement (for example, political appointee, private citizen, defense industry consultant, etc.) The officer's post-retirement occupation is another indicator of how politically active the officer decided to be.

For independent variables, I have to control again for the aggregate levels of institutional military prestige and political polarization, both of which data is available for using Gallup public opinion polling data and DW-NOMINATE scores. But in this chapter, since I am examining the individual officer, I will also control for whether a particular military officer experienced a relative decline, boost, or "no change" to his/her individual status of military prestige while serving in their senior military role. This will require coding for whether the officer experienced reputational damage as a result of an exogenous event, such as: being fired by a President (likely resulting in a loss of prestige), enjoying a tremendous increase in public opinion for directing successful combat operations (likely resulting in an increase in prestige) or, conversely, unsuccessful combat operations (loss of prestige) during war or a major military engagement. Alternatively, I suspect that the vast majority of these senior officers experience neither a boost or decline in individual prestige during their tenure as senior military leaders, i.e. each does his or her duty and then retires. But I suspect that the data will nonetheless include sufficient variation in this variable, over time, that there may be at least a correlation between an officer's individual reputation and their post-retirement political activity.

Obtaining the data for these dependent variables will require classifying what primary role the officer played in retirement (i.e. political appointee, regular civilian, corporate board member), searching the congressional record by the officer's last name and categorizing the number and type of congressional hearings the officer participated in, both while on active duty and during retirement, and searching historical newspaper databases for any articles or speeches the author might have given, both on active duty and during retirement, particularly regarding turbulent or polarizing events (such as implementing the end to segregation (both within the military and in society), national protests during 1968, the response to Hurricane Katrina, etc.). ^[I need to determine which events these are. Similar to the ICB, I have to find the 30 or so "most turbulent domestic events" that have occurred in the United States in the past 70 years, i.e. protests and riots, etc., and focus my research to answer the question, what did these military leaders I am examining have to say about these events? I am scouring the American politics literature for a database or a consensus of what these events are or might be, but as of yet, I have not found one.]

I am not yet clear whether a strict quantitative or qualitative analysis, or a mixture of both, is most appropriate for this chapter. I think that one of the issues this will hinge on is the degree of clarity with which some of the coding decisions can and should be made. In the international relations literature, leader and state reputation is a topic that has been studied, particularly regarding the length of tenure for heads of state as a result of initiation of conflict [@croco_deciders_2011; @croco_war_2016]. What made these studies possible, in my opinion, was clear selection criteria, and particularly the use of data sets involving Militarized Interstate Disputes or International Crises proved helpful. Unfortunately, to my knowledge, there is not a similar data set in the American politics literature that consists of relevant domestic turbulent events of the 20th century, for instance. I am potentially up to the task of building such a list myself, but would require some assistance from an American Politics scholar to do so in a way that passes muster for academic rigor. For example, I can imagine that I would want to include all instances in which military force was used domestically, instances in which there was large, sustained use of the military for responding to domestic natural disasters, and instances in which there was rioting, etc. But there may be more that I need to include, and I would want to bounce this idea off of other scholars to determine how to appropriately selecting the right events.

In chapter five, I plan to present the results of an original survey experiment in which I survey a nationally representative sample of civilians and compare the results to a sample of either a) veterans, or b) active duty military officers (pending a lengthy approval process through the Army, which, quite frankly, may not prove possible given the time constraints I am working with in this particular dissertation). The goal of this survey is to test one of the micro-foundations for the theory that I have laid out, particularly, the role of political polarization in impacting military role beliefs.

The survey will not test the impact of military prestige on military role beliefs, as I assume that everyone taking the survey, in 2020 or 2021, is operating from a standpoint that the current military institution enjoys a relatively high level of prestige and credibility (in other words, the survey assumes that the credibility of the military institution is endogenous). This assumption allows me to focus more on the impact of political polarization on military role beliefs, and to design questions that can further isolate its impact on the dependent variable.

In the survey, the dependent variable I will measure is support for the military engaging in a range of overtly political behaviors in response to controversial situations involving the military. These behaviors span the spectrum of possible outcomes, include obeying the orders, writing an opinion piece, giving an interview, to resignation in protest. I further plan to frame the controversial situations in such a way that each involves the independent variable (IV) of interest, political polarization, through a hypothetical but reasonable (i.e. based on historic events or events that could reasonably occur) situation that is likely to awaken or stimulate two conflicting responses from the survey respondent: a desire to engage in the normatively-correct, non-partisan ethic of the military officer, and a desire to engage in the behavior that is likely affiliated with a particular partisan identity or ideological group. In the series of questions that I plan to ask, half of the population will comprise the treatment group, which will receive an additional piece of information that is highly charged or potentially polarizing that the control group does not receive. I then plan to measure the difference in responses between those whose ideological affiliations are primed (the treatment group) and those whose who were not (the control group), as well as test for differences between the civilian sample and the military/veteran sample. Designing and implementing such a survey will enable me to test two important hypotheses:

_H9: When the salience of one's ideological/partisan identity is raised, there is a greater willingness to engage in overtly political behavior than when one's ideological identity is not raised._

_H10: Relative to civilians, military officers/veterans believe that their identity as a military professional compels them to engage in overtly political behavior when the salience of their ideological/partisan identity is raised._

To test these hypotheses, my survey design will incorporate arguments made by military actors that appeal to the military officer's sense of loyalty. As recent scholarship has argued, the contemporary domestic political environment introduces unique challenges in part because a military officer holds multiple "loyalty structures," or several simultaneous relationships that all thrive in part because a military officer is loyal to the object of these relationships. For example, the military officer is loyal to the President as the Commander in Chief, but also loyal to the Constitution of the United States.

Invoking conceptions of loyalty during the survey is deliberate, and the idea is to appeal to the normative foundation that members of the military are socialized into during their years of service. Each military service emphasizes loyalty as a trait and concept. Of the different military services, the US Army in particular has extensively developed this concept in a framework entitled "The Army Ethic," in which loyalty plays a significant role within the Army profession [@us_department_of_the_army_adp_2019]. This ethic is formally laid out and described in the Army's doctrinal publication entitled, _Army Leadership and the Profession, ADP 6-22_, which charges Army leaders to "embrace and uphold the Army Values and standards of the profession" [@us_department_of_the_army_adp_2019]. There are a total of seven "Army Values," which together spell the acronym, "l-d-r-s-h-i-p," which is short for the word "leadership," and the first of these values, loyalty, is followed by six others: duty, respect, selfless service, honor, integrity, and personal courage [@us_department_of_the_army_adp_2019, p. 1-9.]. By appealing to a military officer's sense of loyalty, the argument made in each question will likely arouse a strong response.

Before asking several questions (likely 3-4), I plan to ask for the respondent's age, sex, partisan affiliation, ideological affiliation, veteran status, and other potential covariates. I will then engage in a series of 4-6 questions, each of which follows the same pattern: presentation of a situation, followed by an elite voice, and an elite opinion. I then will measure the degree to which the respondent agrees with the opinion of the military elite, and the degree to which the respondent agrees with the manner in which the military elite makes his or her opinion known.

Four of these hypothetical situations/questions are as follows:

\singlespace
* Situation #1 (Domestic Use of Force): The President of the United States, a Republican, recently commented that he is seriously contemplating ordering active duty military Troops to assist local police and National Guard units in several US cities in maintaining law and order. Multiple news outlets have shown that these protests, while mostly peaceful, have turned violent in about ten percent of cases over the past several weeks. [Treatment group, add: _These protests are largely affiliated with the Black Lives Matter Organization._]

* Elite Voice: A retired US Army 4 star general, who has extensive combat experience in Panama, Iraq, and Afghanistan, recently penned the following op-ed piece in a prominent newspaper:

* Elite Opinion: “If the President orders the military to deploy to cities where these protests have turned violent, military officers should resign their commissions. Military officers owe a higher loyalty - a sacred loyalty - to the Constitution, which protects the rights of protesters, than they do to any particular President.”

* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)

* To what degree was it appropriate for the retired general to make his opinion public, i.e. to write the op-ed? (7 point scale, strongly disagree - strongly agree)

\doublespace

\singlespace
* Situation #2 (Withdrawal of Troops from Overseas): The President of the United States is considering pulling all US Troops from a middle Eastern country in which American forces have been deployed to for many years as part of the ongoing Global War on Terrorism. Congress is split on this decision. some feel that the President has the authority to make and implement this decision unilaterally, while others have vocally argued that the President should consult Congress before making such an influential decision. Several foreign policy think-tanks have argued that a sudden withdrawal of US forces will result in the harm of innocent civilians by local terrorist groups. [Treatment group, add: _The President is a Democrat and both the House and Senate are controlled by Republicans._]

* Elite Voice: Larry Ferguson is a retired US Navy Admiral, who has extensive overseas experience. He recently gave an interview on a prominent cable news channel in which he stated the following:

* Elite Opinion: “The Chairman of the Joint Chiefs of Staff is the President's chief military adviser. But the Chairman also has an obligation to support and defend the Constitution of the United States, which clearly stipulates that Congress has a duty to exercise an oversight role of the military. If the Chairman does not advise the President to talk to Congress first, he will be failing to uphold his professional duties and responsibilities.”

* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)

* To what degree was it appropriate for the retired general to make his opinion public, i.e. to make these comments during an interview? (7 point scale, strongly disagree - strongly agree)

\doublespace

\singlespace
* Situation #3 (Federal Budget): The President's top economic advisers spent the last several months developing a proposal for next year's federal budget, which was revealed yesterday. Several government agencies experienced funding cuts, to include the Center for Disease Control (CDC), the National Aeronautics and Space Administration (NASA), and the National Academy of Sciences (NAS). [Treatment group, add: _The President has come under fire from some predominantly left-leaning organizations for stating that while he believes in climate change, "the science is not conclusive that it is harming the earth to the degree that some have claimed."_]

* Elite Voice: former Chief of Staff of the Air Force recently gave an interview with NPR radio in which he said, "I truly do not understand why the President would cut the funding from organizations whose job it is to promote science in one form or another. This is par for the course with this President, unfortunately, and it insults the military, too, because we now have to figure out how to work with a President who seems to not value scientific fact. As military professionals, we must uphold facts and seek the truth."

* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)

* To what degree was it appropriate for the retired general to make his opinion public, i.e. to write the op-ed? (7 point scale, strongly disagree - strongly agree)

\doublespace

\singlespace
* Situation #4 (Endorsement of a Presidential Candidate): In the 2024 election, the Republican and Democratic Candidate are very different, particularly regarding their stances on the contentious issue of immigration. [Treatment group, add: _The Democratic candidate has vocally supported a policy of "amnesty" for all immigrants who are currently in the country illegally, while the Republican candidate has stated that, "America is a nation of laws, and without laws, the foundation of society crumbles._"]

* Elite Voice: A retired Marine Corps general spoke at the Democratic National Convention very briefly, introducing the Democratic candidate. He said, "I wrestled with whether I should throw my voice into the political fray, because typically, that's not something we as retired military officers do. But I could not stay silent when I heard the Republican candidate talk about his plans for immigration. And so tonight, I am proud to support the right candidate to be the next President - a candidate who, like me, believes that all people have a right to live in a society that is free from fear. One that is safe, and one that is secure. It is not in spite of the fact that I wore the uniform for 40 years that I make this statement. It is because of it."

* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)

* To what degree was it appropriate for the retired general to make his opinion public, i.e. to speak at the convention? (7 point scale, strongly disagree - strongly agree)

\doublespace

\singlespace
* Situation #5 (Domestic Show of Military Presence/Force): As part of the upcoming Independence Day holiday on July 4th, the President and his administration are mulling the idea of putting on a large parade showcasing the capabilities of the US Military. Some members in Congress are opposed to such a move, as they think it is a political stunt to refocus the nation’s attention away from recent domestic political baggage the President has been dealing with over the past several weeks. [Treatment group, add: The President is a Republican, and the domestic political baggage he has been dealing with refers to new revelations that he and other members of his cabinet may have downplayed the arrival and potential harm caused by a strain of a flu virus that killed more than 10,000 Americans last year.]

* Elite Voice: The Chairman of the Joint Chiefs of Staff, during a news conference at the Pentagon, was recently asked what he felt about the overall situation and the plan to showcase the military. He said: “The military’s duty in this situation is to obey the President, salute the flag, and execute. Period and end of story. There is nothing illegal, immoral, or unethical about what the President wants us to do here, so the military will keep our mouths shut and execute the order.”

* Do you agree with the retired general's opinion? (7 point scale, strongly disagree - strongly agree)

* To what degree was it appropriate for the retired general to make his opinion public, i.e. to speak at the convention? (7 point scale, strongly disagree - strongly agree)

\doublespace

In the sixth chapter, I will turn to a series of qualitative case studies to explore and test the generalizability of my theory. Key in this chapter is finding and examining historical periods in American civil-military relations that are not in the contemporary period, which chapter three and four focus on. I then hope to find evidence for how political polarization and military prestige and credibility impacted the role beliefs adopted by the military during this particular time periods. My goal is to examine all four possible scenarios that Table 2 in the previous pages described, i.e. high and low periods of both polarization and credibility. While I am not settled on the specific historical cases to examine, I have determined an initial breakdown as described in the following paragraphs.

To examine military role beliefs when both political polarization and institutional credibility and prestige are high, I will examine civil-military relations and the role of the military in the post Civil War era. This era is historically fascinating, as it deals with how General Ulysses Grant and other military leaders who enjoyed enormous levels of prestige following the Union's victory in the Civil War, behaved relative to civilian leaders such as President Andrew Johnson during the monumental challenges of reconstruction, and in particular, the Army's role in it, in the southern United States [@hyman_santon_1960].

To examine military role beliefs when polarization is high but credibility and prestige are low, I will focus my attention on the period between roughly 1870 and the start of World War One, or "the gilded age." The military was, at best, in the very beginning of becoming a profession, and the credibility that it had enjoyed following the end of the Civil War had largely if not altogether dissipated. Politically, however, the nation was polarized.

To examine when polarization is low, but the credibility and prestige of the military is high, I may focus on the period after World War Two and the Korean War, from about 1953-1958. Similarly, the period right before World War Two, from about 1933-1940, offers an interesting period of when polarization and military credibility and prestige were both low. A comparison of the military's perceived role in society between these two time periods will help round out the qualitative case study chapter.

In the seventh and final chapter, I will conclude by exploring the implications of my argument.

# Initial Concerns and Timeline

In terms of a plan to finish this dissertation by the desired time of May 2022, I have to set out in earnest.  The following bulletized points capture what I believe are the critical milestones that must be completed.

* Between now and 15 December 2020 - Continue to work on my prospectus. Work with Jonathan Renshon and Jon Pevehouse to develop and field my intended survey experiment. I have already obtained IRB approval through UW-Madison to field a survey, but I need to apply for funding (through the Army), finalize the experiment's questions, and prepare for data collection. Continue collecting data on chapter three.

* January-February 2021 - No later than date to defend my dissertation prospectus and complete my methods comprehensive exam (second field). Continue collecting data for chapter four.

* February 2021 - field my survey experiment (ideally before the inauguration, if possible!).

* March 2021 - Submit chapter three of my dissertation and submit for review to the dissertation committee.

* July 2021 -  Submit chapter four of my dissertation for review to the dissertation committee.

* September 2021 - Submit chapter five of my dissertation for review to the dissertation committee

* January 2022 - Submit chapter six of my dissertation for review to the dissertation committee.

* March 2022 - Submit chapter seven of my dissertation for review to the dissertation committee.

* No later than May 2022 - defend dissertation.

# Conclusion

The purpose of my dissertation is to explore how polarizing domestic politics and increasing military prestige impact the normative conceptions of role beliefs among military leaders. How military leaders view their role within the government will likely impact the degree to which military leaders intervene in politics. My dissertation contends that the two major forces impacting how military leaders view their central roles within government are the degree of political polarization and the level of prestige or credibility of both the military institution and individual military actors. While recent scholars have suspected these forces impact the professional aspects of the military, my dissertation will provide a significant amount of empirical understanding to clarifying the mechanisms through which these trends interact. Moreover, a careful and rigorous analysis will better inform our understanding of how the US military in particular has weathered changes to its role beliefs in the past. Ultimately, this dissertation will yield knowledge that scholars, policymakers, and military officers can each benefit from.

\pagebreak